### 2013-04-10 doc/addresses.tex
Started paper I

\def\oxford{Dept.\ of Physics, University of Oxford, Keble Road, Oxford, OX1 3RH, UK}
\def\kipac{Kavli Institute for Particle Astrophysics and Cosmology, Stanford University, 452 Lomita Mall, Stanford, CA 94035, USA}
\def\ipmu{Kavli Institute IPMU, University of Tokyo, Japan}
\def\adler{Adler Planetarium, Chicago, IL, USA}
\def\zurich{Department of Physics, University of Zurich, Switzerland}

\def\pjmemail{\tt dr.phil.marshall@gmail.com}