### 2014-08-21 doc/acknowledgments.tex

We thank all \Ncollaboration members of the \sw community for their
contributions to the project so far. A complete list of collaborators is
given at... In particular we would like to recognise the efforts of XXX,
YYY etc in moderating the discussion.

We are also grateful to Brooke Simmons, David Hogg, XXX and YYY for many useful
conversations about citizen science and gravitational lens detection. % Need to
cross-ref with Zooniverse author list
%
PJM was given support by the Royal  Society, in the form of a research
fellowship, and by the U.S. Department of Energy under contract number DE-AC02-76SF00515.
% He would like to thank the DotAstronomy community for
% introducing so many valuable concepts and tools, and David Harris and
% Roger Davies for valuable conversations about community-building.
%
% AV \ldots
%
% AM \ldots
The work of AM and SM was supported by World Premier International Research
Center Initiative (WPI Initiative), MEXT, Japan.
%
% Zooniverse / Sloan Foundation.
%
The \sw project is open source, and was developed at
https://github.com/drphilmarshall/SpaceWarps.