### 2015-04-13 doc/acknowledgments.tex
Guessed URL in acknowledgements, for now. Finished going through all AV comments! #96

We thank all \Ncollaboration members of the \sw community for their
contributions to the project so far. A complete list of collaborators is
given at \texttt{http://spacewarps.org/\#/results/CFHTLS}.

We are also grateful to Stuart Lynn, Kelly Borden, Laura Whyte, Brooke Simmons,
David Hogg, Thomas Jennings, Layne  Wright, Cecile Faure, Jonathan Coles, Stuart
Lowe and Jean-Paul Kneib for many useful conversations about citizen science and
gravitational lens detection, and to the Dark Energy Survey and Pan-STARRS strong
lensing science teams for their suggestions and encouragement.

PJM was given support by the Royal Society, in the form of a research
fellowship, and by the U.S. Department of Energy under contract number DE-AC02-76SF00515.
%
AV acknowledges support from the Leverhulme Trust in the form of a research
fellowship.
%
The work of AM and SM was supported by World Premier International Research
Center Initiative (WPI Initiative), MEXT, Japan. The work of AM was also supported in
part by National Science Foundation Grant No. PHYS-1066293 and the hospitality
of the Aspen Center for Physics.
%
% Zooniverse / Sloan Foundation.
%
% Other authors.
PJM and ES thank the Institute of Astronomy and Astrophysics, Academia Sinica
(ASIAA) and Taiwan's Ministry of Science and Technology (MOST) for their
financial support of the workshop ``Citizen Science in Astronomy'' in March
2014, at which some parts of the SWAP analysis was developed.

The \sw project is open source.
The web app was developed at \texttt{https://github.com/Zooniverse/Lens-Zoo}
while the SWAP analysis software was developed at
\texttt{https://github.com/drphilmarshall/SpaceWarps}.