### 2015-04-19 doc/sw-system.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Space Warps Paper I: System
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[useAMS,usenatbib,a4paper]{mn2e}
%% letterpaper
%% a4paper

\voffset=-0.6in

% Packages:
\input psfig.sty
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{epsfig}

% Macros:
\input{macros.tex}
\input{addresses.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title[\SW]
{\SW: Crowd-sourcing the Discovery of Gravitational Lenses}

\author[Marshall et al.]{%
  \input{sw-system-authors.tex}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\date{to be submitted to MNRAS}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}\pubyear{2014}

\maketitle

\label{firstpage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}

We describe \SW, a novel gravitaional lens discovery service that yields samples
of high purity and completeness through crowd-sourced visual inspection.
Carefully produced colour composite images are displayed to volunteers via a
classification interface which records their estimates of the positions of
candidate lensed features. Simulated lenses, and expert-classified non-lenses,
are inserted into the image stream at random intervals; this training set is
used to give the volunteers feedback on their performance, and to estimate a
dynamically-updated probability for any given image to contain a lens. Low
probability systems are retired from the site periodically, concentrating the
sample towards a set of candidates. Having divided the 171 square degree
Canada-France-Hawaii Telescope Legacy Survey (\cfhtls)
imaging survey into some 430,000 overlapping 80 by 80 arcsecond tiles and
displayed them on the site, we were joined by around 37,000 volunteers who
contributed 10 million image classifications over the course of 8 months. The
sample was reduced to 3300 \StageOne candidates; these were then refined to
yield a sample of 90 \StageTwo detections, a sample we expect to be over 90\%
complete and 20\% pure at a threshold of 95\% classification probability. We
comment on the scalability of the \SW system to the wide field survey era.

\end{abstract}

% Full list of options at http://www.journals.uchicago.edu/ApJ/instruct.key.html

\begin{keywords}
  gravitational lensing   --
  methods: statistical    --
  methods: citizen science
\end{keywords}

\setcounter{footnote}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:intro}

% Scientific motivation. Applications of lenses: group-scale arcs,
% galaxy-galaxy lenses, lensed quasars.

Strong gravitational lensing -- the formation of multiple, magnified images of
background objects due to the deflection of light by  massive foreground
objects -- is a very powerful astrophysical tool, enabling a wide range of
science projects. The image separations and distortions provide information
about the mass distribution in the lens \citep[e.g.][]{AugerEtal2010,
SonnenfeldEtal2012,MoreEtal2012,SonnenfeldEtal2013}, including on sub-galactic scales
\citep[e.g.][]{Dalal+Kochanek2002,VegettiEtal2010,HezavehEtal2013}. Any strong
lens can provide magnification of a factor of 10 or more, providing a deeper,
higher resolution view of the distant universe through these ``cosmic
telescopes'' \citep[e.g.][]{StarkEtal2008,NewtonEtal2011}. Lensed quasars
enable cosmography via the time delays between the multiple images'
lightcurves \citep[e.g.][]{TewesEtal2013,SuyuEtal2013}, and study of the
accretion disk itself through the microlensing effect
\citep[e.g.][]{PoindexterEtal2008}. All of these investigations would
benefit from being able to draw from a larger and/or more diverse sample of lenses.

In the last decade the number of these rare cosmic alignments known
has increased by an order of magnitude, thanks to searches carried out in
wide field surveys, such as
CLASS \citep[e.g.]{BrowneEtal2003}, SDSS \citep[e.g.][]
{BoltonEtal2006,AugerEtal2010b,TreuEtal2011,InadaEtal2012,
HennawiEtal2008,BelokurovEtal2009,DiehlEtal2009,FurlanettoEtal2013},
\cfhtls \citep[e.g.][]{MoreEtal2012,GavazziEtal2014}, Herschel
\citep[][]{NegrelloEtal2014} and SPT \citep[e.g.][]{VieiraEtal2013}, among
others.  As the number of known lenses has increased, new types have been
discovered, leading to entirely new investigations. Compound lenses
\citep{GavazziEtal2008,CollettEtal2012} and lensed supernovae
\citep{QuimbyEtal2014,KellyEtal2015} are good examples of this.

% Problem of rarity. Imaging surveys. Problem of purity/false positives.
% Review of progress to date. Methods in SL2S, SQLS. Contrast with SLACS.

Because they are rare, strong lenses are expensive to find. The highest purity
searches to date have made use of relatively clean signals such as the
presence of emission or absorption features at two distinct redshifts in the
same optical spectrum \citep[e.g.][]{BoltonEtal2004}, or the strong
``magnification bias'' towards detecting strongly-lensed sources in the sub-mm
waveband \citep[e.g.][]{NegrelloEtal2010}. Such searches have to yield pure samples,
because they require expensive high resolution imaging follow-up; consequently
they have so far produced yields in the tens to hundreds. An alternative
approach is to search images already of sufficiently high resolution and colour
contrast, and confirm the systems as gravitational lenses by modeling the
survey data themselves \citep[][]{MarshallEtal2009}. Several square degrees of
HST images have been searched, yielding several tens of galaxy-scale lenses
\citep[e.g.][]{MoustakasEtal2007,FaureEtal2008,Jackson2008,MoreEtal2011,
PawaseEtal2014}. Similarly, searches of over a hundred square degrees of CFHT
Legacy Survey (\cfhtls) ground-based imaging, also with sub-arcsecond image quality,
have revealed a smaller number of wider image separation group-scale systems
\citep[e.g.][]{CabanacEtal2007,MoreEtal2012}. Detecting galaxy-scale lenses
from the ground is hard, but feasible albeit with lower efficiency and requiring
HST or spectroscopic follow-up to confirm the candidates as lenses
\citep[e.g.][]{GavazziEtal2014}.

% Scaling to wide field era. Automated methods: problems. Need for good
% training sets. Need for quality control: always present.

How can we scale these lens searches up to imaging surveys covering a hundred
times the sky area, such as the almost-all sky surveys planned with LSST and
Euclid?  There are two approaches to detecting lenses in imaging surveys. The
first one is robotic: automated analysis of the object catalogs and the survey
images. The candidate samples produced by these methods have, to date, not been
of high purity  \citep[see
e.g.][]{MarshallEtal2009,MoreEtal2012,GavazziEtal2014}, with visual inspection
by teams of humans still required to narrow down the robotically-generated
samples. In this approach, the image data may or may not be explicitly modelled
by the robots as if it contained a gravitational lens, but the visual inspection
can be thought of as a ``mental modeling'' step.  An inspector who classifies an
object as a lens candidate does so because the features in the image that they
see can be explained by a model, contained in their brain, of what gravitational
lenses do.  The second approach simply cuts out the robot middleman:
\citet{MoustakasEtal2007,FaureEtal2008,Jackson2008} and \citet{PawaseEtal2014}
all performed successful visual searches for lenses in HST imaging.

Visual image inspection thus seems, at present, unavoidable at some level when
searching for gravitational lenses. The technique has some drawbacks, however.
The first is that humans make mistakes. A solution to
this is for the inspectors to operate in teams, providing multiple
classifications of the same images in order to catch errors and correct them.
Second, and relatedly, is that humans get tired. With a well-designed
classification interface, a human might be able to inspect images at a rate of
one astronomical object per second (provided the majority are indeed
uninteresting). At $10^4$ massive galaxies, and 10 lenses, per square degree,
visual lens searches in good quality imaging data are limited to a few square
degrees per inspector per day (and less, if more time is spent assessing difficult systems).
Scaling to thousands of square degrees
therefore means either robotically reducing the number of targets for
inspection, or increasing the number of inspectors, or both.

For example, a $10^4$ square degree survey containing $10^8$
photometrically-selected massive galaxies and $10^5$ lenses could only be
searched by 10 inspectors at a mean rate of 1 galaxy per second and 10
inspections per galaxy in about 14 years. Reducing the inspection time by a
factor of 400 to two weeks would require a robot to reduce the target sample
to 25 per square degree. However, at this point the required purity, 40\%,
would very likely require the average classification time per object to be
more like 10 seconds per object. Hiring 10 inspectors to assess complex images
full time for five months may not be the most cost-effective or
reliable strategy. Alternatively, a team of $10^6$ inspectors could, in
principle, make the required $10^9$ image classifications, $10^3$ each, in a
few weeks; robotically reducing the target list would lead to a proportional
decrease in the required team size.

Systematic detection of rare astronomical objects by such ``crowd-sourced''
visual inspection has recently been achieved by the online citizen science
project \PH \citep{SchwambEtal2012}.

\PH was designed to enable the discovery of transiting exoplanets in data taken
by the Kepler satellite. A community of over 200,000 inspectors from the general
public found, after each undergoing a small amount of training, over 40 new
exoplanet candidates. They achived this   by visually inspecting the Kepler
lightcurves that were presented in a custom web-based classification interface
\citep{WangEtal2013}.
The older \GZ morphological classification
project \citep{LintottEtal2008} has also enabled the discovery of rare objects,
via its flexible inspection interface and discussion forum
\citep{LintottEtal2009}. Indeed, several of us (AV,CC,CM,EB,PM,TJ,LW) were
active in an informal \GZ gravitational lens search (Verma et al, in
preparation), an experience which led to the present hypothesis that a
systematic online visual lens search could be successful.

In this paper, we describe the \SW project, a web-based system conceived to
address the visual inspection problem in gravitational lens detection for future
large surveys via crowd-sourcing. Implemented as a Zooniverse citizen science
project, it is designed to provide a {\it gravitational lens discovery service}
to survey teams looking for lenses in wide field imaging data. In a companion
paper (More et al, in preparation, herafter \PaperTwo) we will present the new
gravitational lenses discovered in our first lens search, and
begin to investigate the differences between lens detections made in \SW and
those made with automated techniques. Here though, we simply try to answer the
following questions:

\begin{itemize}

\item How reliably can we find gravitational lenses using the \SW
system? What is the completeness of the sample produced likely to be?

\item How noisy is the system? What is the purity of the sample
produced?

\item How quickly can lenses be detected, and non-lenses be rejected?
How many classifications, and so how many volunteers are needed per target?

\item What can we learn about the scalability of the crowd-sourcing approach?

\end{itemize}

Our basic method in this paper is to analyze the performance of the \SW system
on the ``training set'' of simulated lenses and known non-lenses. This allows us
to estimate completeness and purity with respect to gravitational lenses that
have the same properties of the training set. In \PaperTwo, we carry out a
complementary analysis using a sample of ``known'' (reported in the literature)
lenses.

This paper is organised as follows.  In \Sref{sec:design} we introduce the \SW
classification interface and the volunteers who make up the \SW
collaboration,  explain how we use the training images, and describe our
two-stage candidate selection strategy. We then briefly introduce, in
\Sref{sec:data}, the particular dataset used in the first experimental tests
of the \SW system, and how we prepared the images prior to displaying them in
the web interface. In \Sref{sec:swap} we describe our methodology for
interpreting the classifications made by the volunteers, and then present the
results of system performance tests  made on the training images in
\Sref{sec:results}.  We discuss the implications of our results for future
lens searches in \Sref{sec:discuss} and draw conclusions in
\Sref{sec:conclude}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Experiment Design}
\label{sec:design}

The basic steps of a visual search for gravitational lenses are: 1) prepare
the images, 2) display them to an inspector, 3) record the inspector's
classification of each image (as, for example, containing a lens candidate or
not) and 4) analyze that classification along with all others to produce a
final candidate list. We describe step 1 in \Sref{sec:data} and step 4 in
\Sref{sec:swap}. In this section we take a volunteer's eye view and begin by
describing the \SW classification interface, the crowd of volunteers, and the
interactions between the two.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering\includegraphics[width=0.9\linewidth]{sw-system-figs/sw-screengrab-marker+feedback.png}
\caption{Screenshot of the \SW classification interface.}
\label{fig:screenshot}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{The Classification Interface}
\label{sec:design:interface}

A screenshot of the \SW classification interface (CI) is shown in
\Fref{fig:screenshot}. The CI is the centrepiece of the \SW website,
\texttt{http://spacewarps.org}; the web application is written in coffeescript,
css and html and follows the general design of others written by the Zooniverse
team.\footnote{The \SW web application code is open source and can be accessed
from \texttt{https://github.com/zooniverse/Lens-Zoo}. The project was renamed during development.}
The focus of the CI is a
large display of a pre-prepared PNG image of the ``subject'' being inspected.
When the image is clicked on by the volunteer, a marker symbol appears where the
pointer was. Several markers can be placed.

The next image moves rapidly in from a queue formed at the right hand side of
the screen when the ``Finished marking'' button is pressed. Gravitational lenses
are rare: typically, most of the images will not contain a lens candidate, and
these need to be quickly rejected by the inspector. The queue allows several
images to be pre-loaded while the volunteer is classifying the current subject,
and the rapid movement is designed to encourage volunteers to classify rapidly.
After each classification, the positions of the markers are written out to the
classification database, in an entry that also stores the ID of the subject, the
username (or IP address) of the volunteer, a timestamp and some other metadata.

For the more interesting subjects, the CI offers two features that  enable
further investigation of the subjects. The first is the ``Quick Dashboard''
(QD) a more advanced image viewer. This allows the viewer to compare three
different contrast and colour balance settings, to help bring out subtle
features, and to pan and zoom in on interesting regions of the image to assess
small features. Markers can be placed in the Quick Dashboard just the same as
in the main CI image viewer. The second is a link to that subject's page in
the project discussion forum, \texttt{http://talk.spacewarps.org} (known as
``Talk''). Here, volunteers can discuss the features they have seen either
before they submit their classification, or after, if they ``favorite'' the
subject. There is no ``back'' button: each volunteer may only classify a given
subject once. However, the presence of an option to see what others think
about any given subject before submitting your own classification means that
the classifications may not be strictly independent; the advantage of this
system is that volunteers can learn from others what constitutes a good lens
candidate. This is not, however, the primary educational resource; we describe
the explicit training that we provide for the volunteers in the next section.
The FITS images of the subject can be further explored with Zooniverse
``Dashboard Tools,'' using a more powerful image viewer that enables dynamic
variation of the colour balance and contrast (stretch), and for any given view
to be shared via unique URL back to Talk. The FITS image viewing in both the
Quick Dashboard and the Dashboard Tools are enabled by the javascript library
\fitsjs \citep{fitsjs}.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{Training}
\label{sec:design:training}

Gravitational lenses are largely unfamiliar objects to the general public.
New volunteers need to learn what lenses look like as quickly as possible, so
that they can contribute informative classifications. They also need to learn
what lenses do not look like, in order to reduce the false positive detection
rate. There are three primary mechanisms in the \sw system for teaching the
volunteers what to look for. These are, in the order in which they are
encountered, an inline tutorial, instant feedback on ``training images''
inserted into the stream of images presented to them, and a ``Spotter's
Guide.'' As well as this, we provide ``About'', ``FAQ'' and ``Science'' pages
explaining the physics of gravitational lensing. While we expect that the
insight from this static material should help volunteers make sense of the
features in the images, we focus on the more dynamic, activity-based training
early, when engaging new volunteers to participate takes priority.

\subsubsection{Inline Tutorial}

New volunteers are welcomed to the site with a very short tutorial, in which
the task is introduced, a typical image containing a simulated lens is
displayed, and the marking procedure walked through, using pop-up message
boxes. Subsequent images gradually introduce the more advanced features of the
classification interface (the QD and Talk buttons), also using pop-up
messages. The tutorial was purposely kept as short as possible so as to
provide the minimal barrier to entry.

\subsubsection{Training Subjects and Instant Feedback}

The second image viewed after the initial tutorial image is already a survey
image, in order to get the volunteers engaged in the real task as quickly as
possible. However, training then continues, beyond the first image tutorial,
through ``training subjects'' inserted randomly into the stream.  These
training subjects are either simulated lenses (known as ``sims''), or survey
images that were expert-classified and found not to contain any lens
candidates (these images are known as ``duds''). The tutorial explains that
the volunteers  will be shown such training images. They are also informed
that they will receive instant feedback about their performance after
classifying (blind) any of these training subjects. Indeed, after a volunteer
finishes marking a training subject and hits ``Finished marking,'' a pop-up
message is generated, containing either congratulations for a successful
classification (for example, ``Well done! You spotted a simulated lens,'' as
in \Fref{fig:screenshot}) or feedback for an unsuccessful one (for
example, ``There is no gravitational lens in this field!'').

Because we include a training set of images in amongst the survey ``test''
images, \sw can be viewed as a {\it supervised learning} system.
Consequently, we should expect to find lens candidates that look like the sims
that we put in the training set, and so the design of this training set is
quite important. The training images must look realistic, in two ways. First,
apart from the simulated lenses they contain, they must look like all the
other survey images -- otherwise they could be identified as training images
without the volunteer learning anything about what lenses look like. For this
reason, we make the sims by {\it adding in} simulated lensed features to real
images drawn from the survey. It also means that each dataset presented for
inspection in the \sw interface needs to come with its own training set.

Second, the lensed features themselves must be realistic; if they did not
resemble real lenses, we could not expect to find real lenses in the test
images. The details of how we generated the sims for the \cfhtls project are
given in \PaperTwo, where we also carry out performance tests relative to real
lens candidates in the survey fields. Here, we merely note the following
aspects to place this paper's results in context.

We start by selecting massive galaxies and groups in the \cfhtls catalogs,
by colour, brightness and photometric redshift. We then assign plausible mass
distributions to these ``potential lenses:'' a singular isothermal ellipsoid
model for each massive galaxy, and an additional NFW profile dark matter halo
for groups and clusters. The mass distributions are centred, elongated and
aligned with the measured optical properties of the galaxies present. We then
draw a source from a plausible background population, of either galaxies (for
the group or galaxy lenses) or quasars (for the galaxy-scale lenses). The
source redshift, luminosity and size distributions are all chosen to match
those observed, while the source abundance is artificially increased such that
each potential lens has a high probability of having a source occur at a
position that will lead to it being  highly magnified. Source galaxies are
given plausible surface brightness profiles and ellipticities. We then compute
the predicted multiple images of the source using the \gravlens ray-tracing
code \citep{Keeton2000}.

At this point, the distribution of lens and source properties are quite
realistic. We then chose to make cuts of the simulated lens pool, in order to
reject sims that were likely to be diﬃcult to spot. While this ensured the
educational value of the sims when feedback was given, it meant that we should
only expect to detect real lens candidates that also pass these cuts. Any real
lens candidates we ﬁnd with properties outside the ranges of the sims, may in
some sense be considered bonuses, a product of the {\it imagination} of the
crowd in extrapolating their understanding of gravitational lenses into a regime
where they have not seen any before. In the \cfhtls project, the most important
selection cuts were to only keep sims with a)  second brightest image brighter
that $i = 23$ and b) total combined image magnitude brighter than 19--20
(\PaperTwo, Table~1). The combination of these two cuts results in a sample of
visible lensed image configurations.

When a volunteer begins classifying, they are initially shown training images at
a frequency of two in five; these subjects are drawn at random from a pool
containing equal numbers of sims and duds, such that that volunteer never sees a given image more than once. As the number of classifications made by a
volunteer increases, this frequency is decreased, to
$2/(5\times2^{(\textrm{int}(N_c/20)+1)/2})$ ($\approx 0.3$ for the second 20
subjects, $0.2$ for the third 20 subjects, and so on).

This training regime means that in the first 60 images viewed, each volunteer
is shown (on average) 9 simulated gravitational lenses, and 9 empty fields.
This is a much higher rate than the natural one: to try and avoid this leading
to over-optimism among the inspectors (and a resulting high false positive
rate), we display the current ``Simulation Frequency'' on the classification
interface (``1 in 5'' in \Fref{fig:screenshot}) and maintain the consistent
theme in the feedback messages that lenses are rare.

In Figures~\ref{fig:training-gallery:sims}~and~\ref{fig:training-gallery:duds}
we show example training images from the first \SW project (\Sref{sec:data}
below).

\subsubsection{Spotter's Guide}

The instant feedback provides real-time educational responses to the
volunteers as they start classifying; as well as this dynamic system, \SW
provides a static reference work for volunteers to consult when in doubt about
how to perform the task. This ``Spotter's Guide'' is a set of webpages showing
example lenses, both real and simulated, and also some common false positives,
drawn from the pool of survey images. The non-lenses were identified by three
of us (AV, AM and PM) while inspecting a small set of survey images in order
to define the ``dud'' training images. For easy reference, the lenses are
divided by type (for example, ``lensed galaxies,'' ``lensed quasars'' and
``cluster lenses''), as are the false positives (for example, ``Rings and
Spirals,'' ``Mergers,'' ``Artifacts'' and so on). The example images are
accompanied by explanatory text. The Spotter's Guide is reached via a button
on the left hand side, or the hyperlinked thumbnail images of the ``Quick
Reference'' provided on the right hand side, of the classification interface.

Most of the text of the Spotter's Guide focuses on what lenses do or don't do;
the website ``Science page'' contains a very brief introduction to how
gravitational lenses work, which is fleshed out a little on the ``FAQ'' page.
This also contains answers to frequently asked questions about the interface
and the task set.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{Staged Classification}
\label{sec:design:stages}

We now describe briefly the two-stage strategy that was employed in the \cfhtls
project: initial classification (involving the rejection of very large numbers
of non-lenses), and refinement (to further narrow down the sample). The web
application was reconfigured between the two stages, to assist in their
functioning.

\subsubsection{\StageOne: Initial Classification}

The goal of a \StageOne classification is to achieve a high rejection rate,
while maintaining high completeness.  In this mode, therefore, the pre-loading
of images was used to make the sliding in of new subjects happen quickly, to
provide a sense of urgency: initial classification must be done fairly
quickly for the search to be completed within a reasonable time period.  We
expect some trade-off between speed and accuracy, which we return to in the
results section below. Completion of the search requires subjects to be
``retired'' over time, as a result of their being classified. We did this by
analyzing the classifications on a daily basis, as described in
\Sref{sec:swap} below. As subjects were retired, new ones were ingested into
the web app for classification. This means that the discovery of lens
candidates in \StageOne is truly a community effort: to detect a lens candidate,
many non-lenses must first be rejected, and several classifications by different inspectors are needed in either case.

The \StageOne training set was chosen to be quite clear cut, in order to err on
the side of inclusivity, and so ensure high completeness. For the training duds
we selected several hundred images at random, and three of us (AM, with AV and
PM) inspected them and discarded anything that could be considered a lens
candidate: this meant that objects that looks similar to lenses, such as galaxy
mergers, tidal tails and spiral arms, pairs of blue stars and so on, were
specifically excluded from the training set, and so we expect some of those
types of object to appear in the \StageOne candidate list. The training sims for
\StageOne were also selected to be relatively straightforward to spot, as we sought a balance between sensitivity to faint or difficult lenses, and the need to educate the volunteers in what gravitational lenses look like. We return to this balance below.

\subsubsection{\StageTwo: Refinement}

The design of the \StageOne classification task, and its training set, should
lead to sample of lens candidates that has high completeness but may have low
purity. To refine this sample to higher purity, we need to reject more
non-lenses, which means providing the volunteers with a more realistic and
challenging training set as they re-classify it. The more demanding \StageTwo
training set was generated as follows. The \StageTwo duds were selected from the
image pool by at least one of three of us  (AV, AM and PM) as being a potential
false positive, while the \StageTwo sims were chosen to be a subset of \StageOne
sims, none of which were deemed ``obvious'' by the same expert classifiers.
Figures~\ref{fig:training-gallery:sims} and~\ref{fig:training-gallery:duds}
show some example images from the resulting \cfhtls \StageTwo training set.

%%% Sims and Duds from Stage2

\begin{figure*}%[!ht]
\begin{minipage}[b]{0.24\linewidth}
\centering\epsfig{file=sw-system-figs/gallery/4.png,width=\linewidth,angle=0,clip=}
\end{minipage} \hfill
\begin{minipage}[b]{0.24\linewidth}
\centering\epsfig{file=sw-system-figs/gallery/5.png,width=\linewidth,angle=0,clip=}
\end{minipage} \hfill
\begin{minipage}[b]{0.24\linewidth}
\centering\epsfig{file=sw-system-figs/gallery/6.png,width=\linewidth,angle=0,clip=}
\end{minipage} \hfill
\begin{minipage}[b]{0.24\linewidth}
\centering\epsfig{file=sw-system-figs/gallery/7.png,width=\linewidth,angle=0,clip=}
\end{minipage} \hfill
\caption{Typical \SW ``sims'' from the \StageTwo training set. The top right-hand corner insets indicate the
location of the simulated lens in each of these training images. Users needed to
click on these specific features in order to make a correct classification.}
\label{fig:training-gallery:sims}
\end{figure*}

\begin{figure*}%[!ht]
\begin{minipage}[b]{0.24\linewidth}
\centering\epsfig{file=sw-system-figs/gallery/0_ASW0009dgn.png,width=\linewidth,angle=0,clip=}
\end{minipage} \hfill
\begin{minipage}[b]{0.24\linewidth}
\centering\epsfig{file=sw-system-figs/gallery/1_ASW0009dgo.png,width=\linewidth,angle=0,clip=}
\end{minipage} \hfill
\begin{minipage}[b]{0.24\linewidth}
\centering\epsfig{file=sw-system-figs/gallery/2_ASW0009dgy.png,width=\linewidth,angle=0,clip=}
\end{minipage} \hfill
\begin{minipage}[b]{0.24\linewidth}
\centering\epsfig{file=sw-system-figs/gallery/3_ASW0009dh0.png,width=\linewidth,angle=0,clip=}
\end{minipage} \hfill
\caption{Typical \SW ``duds'' from the \StageTwo training set. All of these
subjects were correctly classified by the community as not likely to contain
gravitational lenses.} \label{fig:training-gallery:duds}
\end{figure*}

We also attempted to encourage discernment by changing the look and feel of
the app, slowing down the arrival of new images, and switching the background
colour to bright orange to make it clear that a different task was being set.
The frequency with which training images were shown was fixed at 1 in 3.
Finally, the Spotter's Guide was upgraded to include more examples
of various possible false positives, divided into sub-classes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Data}
\label{sec:data}

We refer the reader to \PaperTwo for the details of the particular set of
imaging survey data used in this first \SW project. Here, we summarize very
briefly the choices that were made, in order to provide the context for our
general description and illustrations of the \SW system.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{The CFHT Legacy Survey}
\label{sec:data:CFHTLS}

The CFHT Legacy Survey\footnote{\texttt{http://www.cfht.hawaii.edu/Science/CFHTLS/}}
\citep[\cfhtls,][]{Gwyn2012} covered 171 square degrees of
approximately equatorial sky, in four patches distributed in right ascension.
With high and homogeneous image quality (the mean seeing in the $g$-band is
0.78''), and reaching limiting magnitudes of around 25 across the $ugriz$ filter
set, this survey has yielded several dozen new gravitational lenses on both
galaxy and group scales
\citep{GavazziEtal2014,SonnenfeldEtal13,CabanacEtal2007,MoreEtal2012}. The
quality of the data, combined with the presence of these comparison ``known
lens'' samples, makes this a natural choice against which to develop and test
the \SW system. The \cfhtls is also well-representative of the data quality
expected from several next-generation sky surveys, such as DES, KiDS, HSC and
LSST.

In order to investigate the completeness of the previous, semi-automated, lens
searches in the \cfhtls area, we designed a ``blind search.'' We divided the
mosaiced images into some 430,000 equal size, overlapping tiles, approximately
80~arcsec on a side. We refer to these images as ``test images'' (as opposed to
``training images''). In future, larger area projects we expect to implement the
rather different strategy of producing image tiles centred on particular
pre-selected ``targets,'' to make for a more efficient (but less complete)
survey. We do not expect the performance of citizen image inspectors to change
significantly between these strategies: to first order, both strategies require
the inspectors to learn what lenses look like, and then search the presented
images for similar features.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{Image Presentation}
\label{sec:data:display}

The \cfhtls $g$, $r$ and $i$-band images have the greatest depth and best image
quality, and we chose to focus on this set (although the $u$ and $z$-band images
were also available for perusal in Talk). We made colour composite PNG
format images following the prescription of \citet{LuptonEtal2004} \citep[with
extensions by][and some particular choices of our own]{WherryEtal2004}, using
the \humvi software.\footnote{The \humvi colour image composition code used in
this work is open source and available from
\texttt{http://github.com/drphilmarshall/HumVI}} Specifically, we first rescaled
the pixel values of each image, in the notation of \citeauthor{LuptonEtal2004},
into flux units (``picomaggies,''), via the image AB zeropoint $m_0$ and the
pixel value calibration factor $f$  given by $\log_{10}{f} = 0.4(30.0 - m_0)$.
We then multiplied these calibrated images by further aesthetic ``scales''
$s_{i,r,g}$ before computing the total intensity image $I$ and applying an
arcsinh stretch. (The scales are re-normalized to sum to one on input.) Thus,
the red ($R$), green ($G$) and blue ($B$) channel images correspond to the
\cfhtls $i$, $r$ and $g$-band images in the following way:
\bea
I & = & (i \cdot s_i + r \cdot s_r + g \cdot s_g), \notag \\
R & = & i \cdot s_i \cdot \frac{{\rm asinh}\left(\alpha \cdot Q \cdot I\right)}{Q \cdot I}, \notag \\
G & = & r \cdot s_r \cdot \frac{{\rm asinh}\left(\alpha \cdot Q \cdot I\right)}{Q \cdot I}, \notag \\
B & = & g \cdot s_g \cdot \frac{{\rm asinh}\left(\alpha \cdot Q \cdot I\right)}{Q \cdot I}. \notag \\
\eea
We chose to allow the composite image formed from these channel images to
saturate to white: any pixels in any of the channel images lying outside the
range 0 to 1 was assigned the value 0 or 1 appropriately. This was not the
recommendation of \citet{LuptonEtal2004}, but we found it to still give very
informative but also familiar-looking astronomical images.

The non-linearity parameters $Q$ and $\alpha$ control the brightness and
contrast of the images. We first tuned $\alpha$ (which acts as an additional
scale factor) until the background noise was just visible. Then we tuned the
colour scales $s$ to find a balance between exposing the low surface
brightness blue features (important for lens spotting!), and having the noise
appear to have equal red, green and blue components. Non-linearity sets in at
about $1/(Q \alpha)$ in the  scaled intensity image: finally, tuning $Q$ at
fixed $\alpha$ determines the appearance of bright galaxies, which we need to
be suppressed enough to allow the low surface brightness features to show
through, but not so much that they no longer looked like massive galaxies.

These parameters were then fixed during the production of all the tiles, in
order to allow straightforward comparison between one image and another, and
for intuition to be built up about the appearance of stars and galaxies across
the survey. Alternative algorithms, such as adjusting the stretch and scale
dynamically according to, for example, the root-mean-square pixel value in
each image, can lead to better presentation of bright objects, but in doing so
they tend to hide the faint features in those images: we needed to optimize
the detectability of these faint features.   Examples of \cfhtls training set
images prepared in this way can be seen in the galleries of the previous
section (Figures~\ref{fig:training-gallery:sims}
and~\ref{fig:training-gallery:duds}). We also defined two alternative sets of
visualization parameters, to display a ``bluer'' image and a ``brighter''
image in the classification interface Quick Dashboard
(\Sref{sec:design:interface}). The QD code performs the same image composition
as just described, but dynamically on FITS images in the browser. These FITS
images are also available for viewing in Talk, via the main Zooniverse
Dashboard image display tool, which again offers the same stretch settings, as
a starting point for image exploration. Our parameter choices are given in
\Tref{tab:data:humvi}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
\begin{center}
\caption{Example \humvi image display parameters, for the \cfhtls images.}
\label{tab:data:humvi}
\begin{tabular}{cccc}
  \hline
  \hline {Image} & Scales ${s_i,s_r,s_g}$ & $\alpha$ & $Q$ \\
  \hline
    ``Standard'' &    ${0.4,0.6,1.7}$     &   0.09   & 1.0 \\
    ``Brighter'' &    ${0.4,0.6,1.7}$     &   0.17   & 1.0 \\
       ``Bluer'' &    ${0.4,0.6,2.5}$     &   0.11   & 2.0 \\
  \hline \hline
\end{tabular}
\medskip\\
\end{center}
\end{table}
% Quck Dashboard code: https://github.com/zooniverse/Lens-Zoo/blob/vics82/app/controllers/quick_dashboard.coffee
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Classification Analysis}
\label{sec:swap}

Having described the classification interface, the training images and the
test images,  we now outline our methodology for interpreting the interactions
of the volunteers with the identification interface, and then describe how we
applied this methodology in the two classification stages in the \cfhtls
project in the \SW Analysis Pipeline (SWAP) code.\footnote{The open source
SWAP code is available from
\texttt{https://github.com/drphilmarshall/SpaceWarps}}

Each classification made is logged in a database, storing subject IDs,
(anonymous) volunteer IDs, a timestamp and the classification results (the image
pixel-coordinate positions of every marker placed).  The ``category'' of subject --
whether it is a ``training'' subject (a simulated lens or a known non-lens) or a
``test'' subject (an unseen image drawn from the survey) -- is also recorded.
For training subjects, we also store the ``kind'' of the subject as
a lens (``sim''), or a non-lens (``dud''), and also the ``flavor'' of lens
object if one is present in the image (``lensed galaxy'', ``lensed quasar'' or
``cluster lens'').  This classification is used to provide instant feedback, but
is also the basic measurement used in a probabilistic classification of every
subject based on all image views to date.

While the \SW web app is live, and classifications are being made, we perform an
online analysis of the classifications,  updating a probabilistic model of every
(anonymous) volunteer's data, and also updating the posterior probability that
each subject (in both the training and test sets) contains a lens, on a daily
basis. This gives us a dynamic estimate of the posterior probability for  any
given  subject being a lens, given all classifications of it to date. Assigning
thresholds in this lens probability then allows us to make good decisions about
whether or not to retire a subject from the system, in order to focus attention
on new images.

The details of how the lens probabilities are calculated are given in
\Aref{appendix:swap}. In summary:
\begin{itemize}

\item Each volunteer is assigned a simple software agent, characterised by a
confusion matrix. The two independent elements of this matrix are the
probabilities, as estimated by the agent, that the volunteer is going to be 1)
correct when they report that an image contains a lens when it really does
contain a lens, $\pr(\saidLENS|\LENS,T)$, and 2) correct when they report that
an image does not contain a lens when it really doesn't contain a lens,
$\pr(\saidNOT|\LENS,T)$.

\item Each agent updates its confusion matrix elements based on the number of
times its volunteer has been right in each way (about both the sims and the
duds) while classifying subjects from the training set, accounting for noise
early on due to small number statistics: $T$ is the set of all training images
seen to date.

\item Each agent uses its confusion matrices to update, via Bayes' theorem,
the probability of an image from the test set containing a lens,
$\pr(\LENS|C,T)$, when that image is classified by its volunteer. ($C$ is the
set of all classifications made of this subject.)

\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering\includegraphics[width=0.9\linewidth]{sw-system-figs/CFHTLS_2013-10-29_10:16:35_probabilities.png}
\caption{Typical \SW agent confusion matrix elements. At a particular
snapshot, 200 randomly-selected agents are shown distributed over the unit plane, with a
tendency to move towards the ``astute'' region in the upper right hand
quadrant as each agent's volunteer views more images. Yellow point size is
proprtional to the number of images classified; green point size shows
agent-perceived ``skill.''}
\label{fig:swap:agent-probabilities}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering\includegraphics[width=0.9\linewidth]{sw-system-figs/CFHTLS_2013-10-29_10:16:35_histories.png}
\caption{Typical \SW agent histories. The ``skill'' of the  same 200 random
agents  as in \Fref{fig:swap:agent-probabilities} is plotted in green as a function
of the number  of subjects classified (``effort''). The yellow histogram in
the background shows the distribution of effort.}
\label{fig:swap:agent-histories}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\Fref{fig:swap:agent-probabilities} shows the confusion matrix elements of 200
randomly-selected agents, as they were on a particular day towards the end of
the \cfhtls project. Many volunteers classify  only a small number of images, and
so their agents' confusion matrix elements remain close to their initial values
of (0.5,0.5). As more images are classified (shown by the yellow point size),
the agents' matrix elements tend to move towards higher values, as the
volunteers attain greater skill levels (green point sizes, see
\Sref{sec:results} below) and the agent learns more about them. In this
quadrant, the agents perceive their volunteers to be ``astute.'' This trend is
more clearly seen in \Fref{fig:swap:agent-histories}, which shows the skill of
the same sample of agents as the number of images classified increases. The
histogram shows the distribution of classification number: a long tail to very
high ``effort'' can be seen.

In \Sref{sec:results} below, we define several quantities based on the
probabilities listed above that serve to quantify the performance of the crowd
in terms of the information they provide via their classifications, and report
on the performance of the system in returning a sample of lens candidates
as a function of $\pr(\LENS|C,T)$ threshold.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering\includegraphics[width=0.9\linewidth]{sw-system-figs/CFHTLS_2013-10-29_10:16:35_trajectories.png}
\caption{Typical \SW \StageOne subject trajectories. Subjects drift downwards in
the top panel as they are classified, while being nudged left and right by the
agents as they interpret the volunteers input. The dotted vertical lines show
(left to right) the retirement threshold, prior probability, and the detection
threshold.}
\label{fig:swap:subject-trajectories}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

During \StageOne classification of the \cfhtls images, we assigned a prior
probability for each image to contain a lens of $2\times10^{-4}$, based on a
rough estimate of the number of expected lenses in the survey, and the fraction
of the survey area covered by each image. We then assigned two values of the
images' posterior probability, $\pr(\LENS|C,T)$, to define ``detection'' and
``rejection'' thresholds. These were set to be 0.95 and (approximately
symmetrically in the logarithm of probability), $10^{-7}$. Subjects that
attained probability of less than the rejection threshold were scheduled for
retirement and subsequently ignored by the analysis code. Subjects crossing the
detection threshold were not retired from the website, but instead left in the
system so that more volunteers could see them (and continue to update their
probabilities). The progress of the subjects is illustrated in
\Fref{fig:swap:subject-trajectories}. Subjects appear on this plot at the tip of
the arrow, at zero classifications and prior probability; they then drift
downwards as they are classified by the crowd, with each agent applying the
appropriate kick in probability based on what it hears its volunteer say.
Encouragingly, sims (blue) tend to end up with high probability, while duds
(red) pile up at low probability; test subjects (black) mostly drift to low
probability, but some go the other way. The latter will help make up the
candidate sample. As this plot shows, around 10 classifications are typically required for a subject to reach the retirement (or detection) threshold.

The analysis code was run every night during the project, and subjects retired
in batches after its completion. This introduced some inefficiency, because
some classifications were accumulated in the time between them crossing the
rejection threshold and the subject actually being retired from the website.
As subjects were retired from the site, more subjects were activated. In this
way, the volunteers who down-voted images for not containing any lensed
features enabled new images to be shown to other members of the community.

When all the subjects had either been retired, or classified around 10 times
or more, the web app was paused and reconfigured for \StageTwo. The sample of
subjects classified during \StageTwo was selected to be all those that passed
the detection threshold ($\pr(\LENS|C,T) > 0.95$) at \StageOne. These were
classified for one week, with no retirement but a maximum classification
number of 50 each. The number of subjects at \StageTwo was small enough that we
did not need to retire any: instead, we simply collected classifications for a
fixed period of time (about 4 weeks). Without the time pressure motivating the
probabilities simultaneously, for comparison. We describe how this differs
from the online analysis in the appendix.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Results}
\label{sec:results}

In this section we present our findings about the performance of the \SW system,
in terms of the classification of the training set, the  information contributed
by the crowd, and the speed at which the image set was classified.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{Sample Properties}
\label{sec:results:sample}

We first quantify the performance of the \SW system in terms of the recovery of
the training set images. At \StageOne, this set contains around 5712 simulated
lenses, and 450 duds; at \StageTwo, we used 152 simulated lenses and 201 duds.
\Fref{fig:results:sample:roc} shows receiver operating characteristic (ROC)
curves for \cfhtls \StageOne and \StageTwo. These plots show the true positive
rate (TPR, the number of sims correctly detected divided by the total number of
sims in the training set), and the false positive rate (FPR, the number of duds
incorrectly detected  divided by the total number of duds in the training set), both
for a given sample of detections defined by a particular probability threshold,
which varies along the curves.  In both stages, these curves show that true
positive rates of  around 90\% were achieved, at very low false positive rates.

For comparison we show the results of an analysis where the classifications of
training images were ignored, and none of the agents allowed to learn: the
agents were instead simply assigned initial values of their confusion matrices
of 0.75 for each element, which then remained constant. This set-up emulates a
very simple unweighted voting scheme, where all classifications are treated
equally.  In this case, the TPR remains under 80\% in \StageOne and 60\% in
\StageTwo, thus quantifying  the benefit of including training images and
allowing the agents to learn via their Bayesian updates.

When the agents are allowed to update their confusion matrices, the choice of
initial confusion matrix is not very important. The same 0.75 initial values
applied to normal, learning agents resulted in only a slightly lower TPR than
the default case, nevertheless indicating that initializing the system less
conservatively results in slightly worse performance.

The dot-dashed curves show the impact of the offline analysis. At \StageOne the
results are very similar to the online version that was actually run (solid
line). However, at \StageTwo there is marginal evidence of there being greater
benefit to doing the analysis offline (\Sref{sec:swap}, \Aref{appendix:swap}).
Over 85\% TPR is achieved at zero FPR in the offline analysis , while if one is
willing to accept a false positive rate of 5\%, the true positive rate rises to
over 95\%, showing that some of the sims that were missed in the online analysis
may be being recovered by doing the analysis offline. The same is true at
\StageOne, but to a lesser extent. Assuming Poisson statistics for the
fluctuations in the numbers of recovered lenses, the uncertainty in the measured
\StageTwo TPR values is around 8\%, but the online and offline samples are
highly correlated, such that the uncertainty on the difference between the ROC
curves is somewhat less than this. Still, a larger validation set is needed to
test these algorithmic choices more rigorously. At \StageOne, high TPR can be
measured to better than 1\%.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\begin{minipage}{0.45\linewidth}
  \centering\includegraphics[width=\linewidth]{sw-system-figs/stage1_ROC.png}
\end{minipage}\hfill
\begin{minipage}{0.45\linewidth}
  \centering\includegraphics[width=\linewidth]{sw-system-figs/stage2_ROC.png}
\end{minipage}
\caption{Receiver operating characteristic curves for the \SW system, using
the \cfhtls training set. Left: \StageOne, right: \StageTwo. Insets show a
magnified view of the top lefthand corner of each plot. Linestyles illustrate
the agent properties (initial confusion matrix elements, learning or not)
discussed in the text, as well as the difference between the online and offline
analyses. The \StageTwo sample was defined using the online \StageOne results
(solid blue curve), while the offline analysis was chosen for the final
\StageTwo results (dot-dashed orange curve). }
\label{fig:results:sample:roc}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Adopting the online \StageOne analysis, and the offline \StageTwo analysis, we
show in \Fref{fig:results:sample:CP} a plot of the more familiar (to
astronomers) quantities of completeness versus purity, again for
the two stages. As in \Fref{fig:results:sample:roc}, the detection threshold
varies along the curves. Completeness is defined as the number of correctly
detected sims divided by the total number of sims in the training set, while
purity is the number of correct detections divided by the total number of
detections.\footnote{The completeness is equivalent to the TPR and is also
known elsewhere as the ``recall.''  The purity is also known as the
``precision.''} If the training set samples from the same systems as the test
set, then the completeness of the training set will be equivalent to the
completeness of the test set. The purity depends on the proportion of sims to
duds, and so the purity of the test set must be approximated by rescaling the
training set to the expected proportion of lens systems to not-lens systems in
the survey. First we compute the expected number of false positives by
multiplying the FPR by the expected number of non-lenses in the survey. Then
we multiply the TPR by the expected number of lenses in the survey, to get the
expected number of true positives. The sum of the true positives and the false
positives gives the expected sample size; dividing the expected number of true
positives by this sample size gives the purity. Note that the completeness is
invariant to this transformation.
The \StageOne curves are truncated by the retirement of subjects in this phase,
which sets the minimum size of this sample. We see from the solid blue curve
that over 90\% completeness was reached, albeit in a sample with around 20\%
purity.

Does this performance level vary between the different types of gravitational
lens? To investigate the completeness to the three different types of lens in
the training set, we repeated the same procedure but now considering only the
detections of a certain kind of lens and of the non-lenses in the training set.
We estimate the expected number of lenses and non-lens false positives by
dividing the lens and dud sets into equal fractions. The lensed quasar part of
the training set yielded the highest completeness, suggesting that these were
the easiest sims to spot. The lensed galaxies were recovered at the lowest
completeness, likely due to the difficulty  of separating the lens and source
galaxy light.

At \StageTwo, where no retirement was carried out, it was possible to reach
100\% purity: the knee of the curve is at just under 90\% completeness. However,
the purity decreases rapidly if higher completeness than this is sought. The
optimal sample in this simulated lens search experiment would have been
constructed with a threshold value of $\pr(\LENS|C,T) > 0.47$. At 100\% purity
and 90\% completeness, it would have contained around 89 lens candidates.

% TP + FN ~ 2e-4 * 430000 ~ 100 ~ TP/C ; so TP ~ 89 candidates

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering\includegraphics[width=\linewidth]{sw-system-figs/completeness_purity_simple.png}
\caption{Completeness-estimated purity curves for the \sw system, using
the \cfhtls training set.}
\label{fig:results:sample:CP}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{Crowd Properties}
\label{sec:results:crowd}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering\includegraphics[width=0.9\linewidth]{sw-system-figs/all_skill_contribution_experience_education.png}
\caption{Key properties and contributions of the \SW crowd. Plotted are the 1-D
and 2-D smoothed marginalized distributions for  the logarithms of the
properties of the agents described in the text. The  \StageOne agents are shown
in blue, the \StageTwo agents in orange: the  smoothing scale for the \StageTwo
curves was set slightly higher to counter the  higher shot noise. The contours
contain 68\%, 95\% and 99\% of the distributions.}
\label{fig:crowd:cornerplot}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To investigate the properties of the \sw crowd, as characterised by their
agents, we define the following quantities and plot their distributions in
\Fref{fig:crowd:cornerplot}.
%
\begin{description}
%
\item{\noindent\bf ``Effort:''} The number of test images, $\effort$, classified
by a volunteer. In \StageOne, the mean effort per agent was 263; in the shorter
\StageTwo it was 81.
%
\item{\noindent\bf ``Experience:''} The number of training images, $\experience$,
classified by a volunteer. In \StageOne, the mean experience per agent was 29;
in \StageTwo (where the training image frequency was set higher) it was 34.
%
\item{\noindent\bf ``Skill:''} The expectation value of the information gained
per classification (in bits) by that volunteer,  $\skill$ for subjects which
have lens probability 0.5 (\Aref{appendix:swap}). Random classifiers have
$\skill = 0.0$, while perfect classifiers have $\skill = 1.0$. All agents start
with $\skill = 0.0$. The skill of  an agent increases as training subjects are
classified, and the agent's estimates of its confusion matrix elements improve.
In \StageOne, the mean skill per agent was 0.04 bits; in  \StageTwo it was 0.05.
%
\item{\noindent\bf ``Contribution:''} The integrated skill over a volunteer's
{\it test subject} classification history, a quantity representative of the
total contribution to the project made by that volunteer (see the appendix for
more discussion of this quantity). The classifications of training images
determine the skill, while the classification of test images determines
contribution.  In \StageOne, the mean  contribution per agent was
34.9 bits; in \StageTwo it was 33.5.
%
\item{\noindent\bf ``Information:''} The actual total information $\information$
generated by the agent during the volunteer's classification activity. This
quantity depends on the value of each subject's lens probability when that
subject was presented to the volunteer (simply because information gain is
defined in terms of posterior relative to prior probabilities,
\Aref{appendix:swap}).  Because of this, we expect this quantity to be noisier
than the ``contribution'' defined above.
%
\end{description}

The leftmost column of \Fref{fig:crowd:cornerplot} shows how the last four of
these properties depends on the effort expended by the volunteers. As expected,
we see that experience is tightly correlated with effort, reflecting the design
of training images being presented throughout each stage (albeit at decreasing
frequency). We also see a strong correlation between effort and skill, which was
hoped for but not guaranteed: the more images the volunteers see, the better
able to contribute information they are.

The effort distribution shows two peaks, suggestive of two types of
participation.  The sharp spike at just a few images classified presumably
corresponds to visitors who only classify a few subjects before leaving the site
again. The broader hump contains people doing tens to hundreds of
classifications: the skill vs.\ experience panel of \Fref{fig:crowd:cornerplot}
shows this group to achieve a broad distribution of  skill, peaking at around
0.05 but with a long tail to higher values.

At high values of experience and effort, the skill is \emph{always high}. There
seem to be very few agents logging large numbers of classifications at low skill
(although there are one or two exceptions): almost all high effort
``super-users'' have high skill. These two properties are reflected in both the
contributions these volunteers make (third row) and the information they
generate (fourth row).

The distributions for the \StageTwo agents (orange) are qualitatively similar to
those for the \StageOne agents (blue). Differences are: 1) the maximum effort
possible at \StageTwo is smaller, because fewer subjects were available to be
classifed, but 2) the mean effort expended at \StageTwo was higher (perhaps
because the subjects were higher probability, and more compelling); 3) the
information generated per agent was higher at \StageTwo, because the subjects
had higher probability.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering\includegraphics[width=0.9\linewidth]{sw-system-figs/crowd_contrib_cumul.png}
\caption{Cumulative distributions of the contributions made by the agents: 90\%
of the contribution was made by the highest contributing 1\% of the crowd at
\StageOne (blue), and the highest contributing  7\% of the crowd at \StageTwo
(orange). The \StageOne agents are shown in blue, the \StageTwo agents in
orange. ``Experienced volunteers'' classified 10 or more training subjects.}
\label{fig:crowd:cumulplot}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\Fref{fig:crowd:cornerplot} shows the \SW crowd to have quite broad
distributions of logarithmic effort, skill, and contribution. To better
quantify the contributions made by the volunteers, we show their cumulative
distribution on a linear scale in \Fref{fig:crowd:cumulplot}. This plot shows
clearly the importance of the hardest-working, most active volunteers: at
\StageOne,  1.0\% of the volunteers -- 375 people -- made 90\% of the
contribution.  At \StageTwo, where it was not possible to make as
many classifications before running out of subjects, 7.2\% of the volunteers
-- 141 people -- made 90\% of the contribution.

However, it is not the case that only these small groups were {\it capable} of
making large contributions. The cumulative distribution of agent skill is shown
in \Fref{fig:crowd:cumulskillplot}: these distributions are significantly
broader than the corresponding distributions of agent contribution, in
\Fref{fig:crowd:cumulplot}. The most skilled 20\% of agents only possess around
80\% of the skill at each stage. The inexperienced volunteers also possess a
significant fraction of the skill: the most skillful 20\% of experienced
volunteers (1824 people)  possess just 43\% of the total skill. The breadth of
the skill distributions suggests that  the high level of contribution made at
\SW by experienced volunteers is largely a matter of choice (or perhaps,
availability of time!). There were many other volunteers that  were skillful
enough to make large contributions -- they just didn't classify as many images.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering\includegraphics[width=0.9\linewidth]{sw-system-figs/crowd_skill_cumul.png}
\caption{Broad cumulative distributions of agent skill:
the most skilled 20\% of the crowd only possess 79\% of the skill at \StageOne.
The \StageOne agents
are shown in blue, the \StageTwo agents in orange. ``Experienced volunteers''
classified 10 or more training subjects.}
\label{fig:crowd:cumulskillplot}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{Classification Speed}
\label{sec:results:speed}

How fast does the \SW crowd classify subjects? Each agent records the timestamp
of each classification its volunteer makes; by measuring the time lags between
successive classifications we can make estimates of the crowd's classification
and contribution speed. We plot these two quantities for both \StageOne and
\StageTwo in \Fref{fig:results:speed}, normalizing the speed and time axes to
the respective totals. The fractional classification rates in each stage of the
survey fall off in approximately the same way, despite the factor of nearly 20
difference in crowd size. \StageOne (consisting of $\sim430,000$ subjects) was
completed in around 5000 hours (by some 37,000 participants), with
classification rates in the first few days averaging at $\sim10^4$ per hour.
This was stimulated by various forms of advertising (press releases, and emails
to registered Zooniverse users). The asymptotic classification rate was  around
ten times lower, at $\sim10^3$ per hour. Interestingly, the average skill was
found to be approximately constant over the lifetime of \StageOne, leading to a
contribution rate that tracks closely the classification rate. (A 10\% increase
in average skill was seen over the first 40 days, and a decrease of the same
amount in the last 40 days -- not enough to cause a significant difference
between classification and contribution rate behavior, but perhaps reflecting a
learning period at the start and a decrease in participation late on.) \StageTwo
(3700 subjects) was completed in around 600 hours by around 2000 participants,
with classification rates starting at $~\sim2000$ per hour and decaying to
$\sim100$ per hour.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering\includegraphics[width=\linewidth]{sw-system-figs/rate_classification.png}
\caption{Fractional classification rate in the \SW \cfhtls survey.}
\label{fig:results:speed}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The similarities between \StageOne and 2, suggest that these numbers could be
scaled to estimate the speed of future \SW projects (or at least those following
a similar blind search strategy). For example, the completion time for a \SW
project could be approximated as
\begin{equation}
    \tau \approx 18\,{\rm days} \left(\frac{10^4}{N_p}\right)
                               \left(\frac{N_s}{10^4}\right),
\end{equation}
where $N_s$ is the number of subjects to be classified, and $N_p$ is the number
of volunteers in the crowd.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Discussion}
\label{sec:discuss}

What can we learn from the results of the previous section, for future
projects? Potential improvements to the \SW system can be divided into three
categories, performance, efficiency and capacity.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\begin{minipage}{\linewidth}
  \begin{minipage}[t]{0.47\linewidth}
    \begin{minipage}{0.46\linewidth}
      \centering\includegraphics[width=\linewidth]{sw-system-figs/ASW0008wyp_gri.png}
    \end{minipage}\hfill
    \begin{minipage}{0.50\linewidth}
      \centering\includegraphics[width=\linewidth]{sw-system-figs/ASW0008wyp_stage2_trajectory.png}
    \end{minipage}
  \end{minipage}\hfill
  \begin{minipage}[t]{0.47\linewidth}
    \begin{minipage}{0.46\linewidth}
      \centering\includegraphics[width=\linewidth]{sw-system-figs/ASW0006rkx_gri.png}
    \end{minipage}\hfill
    \begin{minipage}{0.50\linewidth}
      \centering\includegraphics[width=\linewidth]{sw-system-figs/ASW0006rkx_stage1_trajectory.png}
    \end{minipage}
  \end{minipage}
\end{minipage}

\vspace{\baselineskip}

\begin{minipage}{\linewidth}
  \begin{minipage}[t]{0.47\linewidth}
    \begin{minipage}{0.46\linewidth}
      \centering\includegraphics[width=\linewidth]{sw-system-figs/ASW0000fgj_gri.png}
    \end{minipage}\hfill
    \begin{minipage}{0.50\linewidth}
      \centering\includegraphics[width=\linewidth]{sw-system-figs/ASW0000fgj_stage1_trajectory.png}
    \end{minipage}
  \end{minipage}\hfill
  \begin{minipage}[t]{0.47\linewidth}
    \begin{minipage}{0.46\linewidth}
      \centering\includegraphics[width=\linewidth]{sw-system-figs/ASW0003j2s_gri.png}
    \end{minipage}\hfill
    \begin{minipage}{0.50\linewidth}
      \centering\includegraphics[width=\linewidth]{sw-system-figs/ASW0003j2s_stage1_trajectory.png}
    \end{minipage}
  \end{minipage}
\end{minipage}
\caption{Illustrative examples of
false negatives (blue) and false positives (red) from the classification of
the \SW \cfhtls training set. The trajectory plot for each of the 4 subjects is
shown to the right of its image. Lefthand panels show the most common types of
trajectories: short step random walks resulting from no high skill classifiers
viewing the subject. The righthand panels show examples where higher skill
classifiers have been involved, causing random walks with longer step lengths
(top) and catastrophic mis-classifications (bottom right). Note that both these
subjects were {\it very nearly} detected but didn't quite reach the threshold
set (marked by the blue dashed line). Insets for the false
negatives show the lens feature. The inset for the lower righthand panel shows
where most users believed a lens was located. There is no corresponding inset
for the lower left panel because no particular feature stood out to users.
Instead, users clicked many different features.}
\label{fig:discuss:performance:trajectories}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Improving performance: reducing incompleteness and impurity}
\label{sec:discuss:performance}

We investigated the source of the incompleteness and impurity visible in
\Fref{fig:results:sample:CP}, by examining the \StageTwo false negatives and
false positives, and their behaviour as they are classified using the online
analysis trajectory plots introduced in \Sref{sec:swap}
(\Fref{fig:swap:subject-trajectories}).
\Fref{fig:discuss:performance:trajectories} shows 2 example simulated lenses
that were missed ($\pr(\LENS|C,T) < 10^{-5}$) by the \SW system (top row),
and 2 example
non-lenses that were incorrectly flagged as candidates ($\pr(\LENS|C,T) >
0.95$) by the \SW system (bottom row).
While the final \StageTwo analysis was done offline,
these trajectory plots still serve to illustrate the approximate
classification histories of the subjects.

In some cases the rejection of the false negatives is understandable: the
lensed features are faint, or in some cases, somewhat unrealistic as lenses.
However, in other cases a reasonably obvious lens was passed over.  This
mainly seems to be due to noise in the system: when only low-skill classifiers
view a subject, all the updates to its posterior probability are small, and if
none are very confident about the presence of lensing, the subject follows a
random walk down its trajectory plot. This can be seen in the top left panel
of the figure.

The false positives show similar behaviour, for example in the bottom lefthand
panel of the figure. As well as subjects being ``unlucky'' in this way, there
are two less common failure modes associated with mistakes made by higher skill
users, illustrated in the righthand column of the figure. In the false negative
subject shown in the top right-hand panel, several high skill classifiers update
the subject upwards in log probability by some way each time, but other,
comparable skill classifiers mis-classify the system to lower probability.   The
trajectory looks like a random walk, but with bigger steps; this particular
subject came very close to crossing the detection  threshold three times, but
didn't quite make it. The bottom right-hand panel shows an example of a final,
apparently rare, failure mode: we see some short-step random walk behaviour,
followed by a mis-classification by a very high skill ``expert'' classifier
after 20 classifications that promotes the subject to high probability.

There are a number of places where we can address these problems and
improve system performance: adding flexibility to the classification
interface, educating the volunteers, assigning subjects for
classification, and interpreting the classification data.

Some of  the mistakes made by reasonably high skill
classifiers working at high speed
could have perhaps been corrected by those classifiers themselves, had
they had access to a ``go back'' option. While clearly enabling error
reduction, we might worry about such a mechanism enabling classifiers to have
a guess, then go and read the Talk discussion, and then return to the
classification interface to change their mind to agree with the forum. In
fact, this is already possible on the current interface -- and indeed, it is
not obvious that such correlations would necessarily lead to poorer system
performance. This could be tested by presenting a fraction of the volunteers
with a version of the site that actively suggested that they take this
approach, and then tracking the relative performance of ``collaborative''
classifications compared with independent ones. We leave the exploration of
this to further
work.

Mistakes by both high and low skill classifiers could be reduced by improving
the training in the system, which could be done in several ways. One is to
make more training images available to those who want or need it. A basic
level of training images are needed for SWAP to build up an accurate picture
of each classifier's skill -- but one could imagine volunteers {\it choosing}
to see more training images (still at random) in their stream. We could also
experiment with providing greater training rates early on for all volunteers;
the risk associated with this is that retention rates may drop if too few
``fresh'' test images are shown early on. Another way of improving the
training could be to provide more information about what lenses do. In this
project, the Spotter's Guide, and the Science, FAQ and About pages were always available on the site, but as a
passive background resource. We might consider providing more links to this
guide in the feedback messages shown to the classifiers as they go -- or
perhaps extending these messages to themselves include more explanations and
example images. We might also investigate a more dynamic Spotter's Guide: a
set of manipulable model lenses illustrating all the possible image
configurations that that those deflectors can make could help volunteers gain
understanding of what lensed features can look like very quickly. Such a toy
is under development.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering\includegraphics[width=0.9\linewidth]{sw-system-figs/stage2_veteran_contribution.png}
\caption{\StageTwo classifier skill, as a function of their \StageOne skill.
Veterans from \StageOne  are shown in blue, while new volunteers are shown in
orange. Point size is proportional to the number of \StageTwo classifications
made, while dashed lines are drawn at the mean values for each sample.
Restricting \StageTwo to classifiers with high \StageOne skill could reduce the
noise in the system.}
\label{fig:discuss:performance:stage1vstage2}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

However, most identification failures (around 60\%) seem to be due to the
random walk problem. Short, random direction steps arise from low skill
classifiers, arising from classifier inexperience (\Sref{sec:results:crowd}).
While improved training will help reduce this effect, it is here
that targeted task assignment could also help improve system performance, by
bringing higher skill volunteers in to the mix. We
advertised \StageTwo of the current project to all registered users; it was
taken up by \StageOne veterans with a broad range of skill, and also picked up a
significant number of new users who did not have enough time to aquire high
skill (since \StageTwo was quite short): of the 1964 volunteers who took part in
the \StageTwo classification round, only 774 were veterans from \StageOne.
\Fref{fig:discuss:performance:stage1vstage2} shows how the \StageOne skill maps
on to \StageTwo skill and contribution.

This figure suggests that there may be some benefit to opening \StageTwo on the
basis of \StageOne agent skill, in order to reduce the noise in the system
generated by new and low skill volunteers in this more difficult
classification stage.  A simple implementation of this would be to switch the
web app to the \StageTwo version when either a certain \StageOne skill (and/or
effort) threshold was passed. Good strategy would probably be to switch back
to \StageOne after some number of \StageTwo classifications (to avoid losing the
high skill users from \StageOne entirely), and also to make it very clear which
stage the classifier was at (since \StageTwo requires more careful attention).
This is something we plan to experiment with in future.

Finally, there could be some performance gains to be made by improving the
SWAP agent model, or its implementation. Low skill is typically a result  of
inexperience (\Sref{sec:results:crowd}), but it could be the agent that is
inexperienced as well as the classifier or its agent. In a future paper we
plan to investigate the use of the test images as well as the training images
in accelerating the agent's learning (Davis et al, in preparation). We also
plan to investigate the use of offline analysis at \StageOne, for the same
reason (see also \Sref{sec:discuss:efficiency} below.)
One might also consider looking at
introducing more conditional dependences in the confusion matrix elements, to
allow for some classifiers having greater skill in spotting one type of lens
than another, or, more generally, as a function of lens property (such as
colour, brightness, and image separation). In the current model, all agents are
considered to be completely independent, whereas in fact we might expect there
to be significant clustering of the confusion matrix elements. A hierarchical
model for the crowd, with hyper-parameters describing the distribution of
confusion matrix elements across the population, may well accelerate the agent
learning process by including the notion of one agent being likely to be
similar to its neighbours. Finally, it is worth noting that the model of
\citet{IBCC} explicitly avoids the assumption of the agent confusion matrix
elements being constant in time, allowing the development of volunteer skill
to be more accurately tracked -- and that they did see some time-evolution in
the supernova zoo classifiers' skill. Finding a way to incorporate such a
learning model into SWAP, while retaining its statistically online character
is an interesting challenge for future work.

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \begin{figure}
% \centering\includegraphics[width=0.9\linewidth]{sw-system-figs/early_vs_final_skill.png}
% \caption{Predictability of final skill from early skill (after 10 training
% classifications) in \StageOne agents. The dashed lines enclose 649 agents with
% early skill greater than 0.1
% who retain final skill of greater than 0.05. Point size is proportional to
% contribution (plus a constant).}
% \label{fig:crowd:skillprediction}
% \end{figure}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Having seen the importance of the high skill volunteers at both \StageOne and
% \StageTwo, we might ask, can we predict agent skill after, say, the volunteer
% has experienced 10 training subjects? In \Fref{fig:crowd:skillprediction} we
% plot this ``early skill'' against final skill, for all experienced \StageOne
% volunteers. The majority of the crowd shows little correlation, but at the
% high early skill end, some predictability appears. The 649 experienced \StageOne
% volunteers who had early skill greater than 0.1 went on to attain a mean
% final skill of 0.22, with 97\% remaining at skill 0.05 or higher. This
% suggests that it might be worth tracking volunteers' skill as a project
% progresses, in order to encourage those showing an aptitude for the task to
% take part in the more difficult activities, such as \StageTwo refinement.
% However, this would not have done a very good job at identifying the
% volunteers who went on to make the largest contributions: the point size in
% \Fref{fig:crowd:skillprediction} is proportional to contribution (plus a constant):
% many of the highest contributors showed relatively low skill early on.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{Improving efficiency}
\label{sec:discuss:efficiency}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}
\begin{center}
\caption{Total crowd and subject sample properties from the \cfhtls project.}
\label{tab:crowd:contributions}
\begin{tabular}{cccccccc}
  \hline
  % \hline {Stage} & Subjects & Contribution                          & Agents & Skill      & Classifications          & Candidates & Information \\
  %                & $\Ns$    & $\sum_k^{\Nv} \contribution_k$ (bits) & $\Nv$  & $\sum_k^{\Nv} \skill_k$ (bits) & $\sum_k^{\Nv} \thiseffort$ & $\Ncands$  & $\sum_j^{\Ns}\sum_k^{\Nv} \information_{j,k}$ (bits) \\
  % \hline
  %           1    & $427064$ & $1292016.3$ & $36982$ & $1471.9$ & $10802125$ & $3368$ & $91122.6$ \\
  %           2    & $3679$   &   $21895.8$ &  $1964$ &  $102.4$ &   $224745$ &   $90$ &  $1640.4$ \\
  \hline {Stage} & Subjects & Contribution                          & Agents & Skill      & Classifications          & Candidates & Information \\
                           & $\Ns$    & $\sum_k^{\Nv} \contribution_k$ (bits) & $\Nv$  & $\sum_k^{\Nv} \skill_k$ (bits) & $\sum_k^{\Nv} \thiseffort$ & $\Ncands$  & $\sum_j^{\Ns}\sum_k^{\Nv} \information_{j,k}$ (bits) \\
            \hline
                      1    & $427064$ & $1292016.3$ & $36982$ & $1471.9$ & $10802125$ & $3368$ & $91122.6$ \\
                      2    & $3679$   &   $21895.8$ &  $1964$ &  $102.4$ &   $224745$ &   $90$ &  $1640.4$ \\
  \hline \hline
\end{tabular}
\medskip\\
\end{center}
\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\Tref{tab:crowd:contributions} shows the total effort, contribution, skill and
information generated in both \StageOne and 2 of the \cfhtls project, with the
total numbers of agents and subjects for comparison.   These numbers allow us
to quantify the efficiency of the system.

The contribution per classification is defined in terms of a hypothetical
subject with lens probability of 0.5; one bit of information is needed to
update such a subject's lens probability to either zero or one. This means
that a maximally complete classification stage would yield a total
contribution (summed over all agents) equal to the number of subjects. The
ratio of this hypothetical optimum to the actual total contribution is
therefore a measure of the stage's inefficiency. We find our inefficiency (by
dividing column 2 by column 3) to be 33\% and 17\% in \StageOne and \StageTwo
respectively. In \StageOne, this inefficiency is due to the daily processing: we
were not able to retire subjects fast enough, and so they, remained in the
system, being over-classified. Indeed, only 3705745 classifications were
needed to retire all the subjects: the ratio of this to the total number made
is 34\%. (The remaining 1\% is due to not all subjects being classified to 1
or 0 probability.) At \StageTwo, we did not retire any subjects at all; the
inefficiency in this case was by design, to give everyone a chance to
appreciate what they had found together. (An unwanted side effect of this
policy was noted in \Fref{fig:discuss:performance:stage1vstage2} in the
previous sub-section.)

It is clear that to increase the efficiency of the system we need to reduce
the time lag between the classification being made and its outcome being
analyzed. The optimal way to do this would be to have the web app itself
analyze the classifications in a fully online system. This is under
investigation for future projects. With the classification data being analyzed
in real time, there may still be a place for a daily or weekly offline
analysis: this could potentially reduce the false negative and false positive
rates by ``resurrecting'' subjects that had been retired by the online system
before the agents had time to learn enough about their classifier's high
skill.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{Increasing capacity}
\label{sec:discuss:capacity}

Finally, we comment on the size of the first \SW crowd, and how the system
might be scaled up for future surveys. In \Sref{sec:results:crowd} and
\Tref{tab:crowd:contributions}, the following rough picture emerged for the
\SW crowd in this project:
it consisted of a few $10^4$ volunteers, with a few $10^3$ achieving
considerable skill, and a few $10^2$ having the time to make a significant
contribution. Slightly more quantitively, we might note that  the total skill
of the crowd, computed by summing the skill of all the agents, is a measure of
the effective crowd size, in the sense that a crowd of perfect classifiers
would be of this size. By this measure,the \StageOne crowd was equivalent to a
team of 1470 perfect classifiers, while the \StageTwo crowd was equivalent to a
team of 102 perfect classifiers. With this same crowd, we can see that surveys
providing a few $10^4$ subjects would be completed quickly, if the high
contribution rate of the current crowd were to be repeated.

There are (at least) two ways in which we might increase the numbers of
high-contribution volunteers for larger projects in future. The first is
simply to increase the total crowd size, and hope that a similar fraction of
volunteers make large contributions. Greater exposure of the website to the
public through mass media would help. Another option is to advertise the
project to new groups of volunteers by translating the website into other
languages (something which is now supported by the Zooniverse). A
multi-lingual user base would come with its own set of challenges, especially
in terms of volunteers' continuing training and interactions on Talk.

The second way to scale up the number of high contribution classifiers is to
increase the rate at which new volunteers become dedicated volunteers. Based
on feedback from the wider \SW community, this could potentially be achieved
through closer collaboration with the science team. It's also possible that
the dynamic \StageTwo system proposed in the previous section may act as an
incentive to some volunteers to increase their contribution (while gaining
skill in order to ``reach'' \StageTwo). Reducing the rate at which  new
volunteers lose interest could also play a role. Anecdotally, it seems fairly
common for new volunteers to be wary of classifying at all, for fear of
introducing errors. Better explanation of how their early classifications are
analyzed could help assuage these fears: \Sref{sec:results:crowd} shows that
effectively down-weighting new users' classifications (by setting their
agents' initial confusion matrix elements to those of a random classifier)
leads to best performance, a result which should be of some comfort to the
nervous volunteer.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions}
\label{sec:conclude}

We have designed, implemented and tested a system for   detecting new strong
gravitational lens candidates in wide field imaging surveys by crowd-sourced
visual inspection. The \SW web-based classification interface presents carefully
prepared colour composite sky images to volunteers, who mark features they
propose to  have been lensed. The participants receive ongoing training with a
mixture of simulated lens and known-to-be-empty images, and we use this
information to inform the automated interpretation of their markers. In our
first lens search we simply divided the \cfhtls imaging into some 430,000 tiles,
and collected over 10~million volunteer image classifications.

% Questions we set ourselves in the introduction:
% \begin{itemize}
% \item How reliably can we find gravitational lenses using the \SW
% system? What is the completeness of the sample produced likely to be?
% \item How noisy is the system? What is the purity of the sample
% produced?
% \item How quickly can lenses be detected, and non-lenses be rejected?
% How many classifications, and so how many volunteers are needed per target?
% \item What can we learn about the scalability of the crowd-sourcing approach?
% \end{itemize}

By analyzing the  classifications made of the training set, we conclude that
gravitational lens detection by crowd-sourced visual inspection works,  and in
the following specific ways:

\begin{itemize}

\item Participation levels were high (over 36,000 volunteers, contributing
classifications at rates between $10^3$ and $10^4$ images per hour), suggesting
that if this can be maintained, visual inspection of tens of thousands of images
can be performed on timescales of a few weeks.

\item Over the course of the two stages of the \cfhtls project, the
set of images was reduced by a factor of $10^3$ (from a few hundred thousand to
a few hundred), enough for an expert team to take over.

\item The ``skill'' of the volunteers correlates strongly with the number of
images they have classified.  Quantifying the ``contribution'' as the integrated
skill over the test subjects  classified, we find that about 90\% of the
contribution was made by a few  hundred volunteers -- but that the width of the
skill distribution means that a  few thousand volunteers could have made
comparable contributions, had they classified more images.

\item The optimal true positive rate (completeness) and false positive rate in
the training set were estimated to be around $91--93\%$ and $<1\%$ (Stages 1 and
2). This  FPR translates to a purity of 20\%.

\item The simulated gravitational lenses that were missed were predominantly
galaxy-scale lenses with faint blue galaxy sources, whose lensed features are
difficult to distinguish from the light from the lens galaxy. We observed some
additional scatter, with some simulated lenses accumulating low probability
simply due to classification noise.

\end{itemize}

In \PaperTwo we present the results of this \cfhtls lens search in more detail,
in particular focusing on the detection of real lenses, and the comparison
between \SW and various automated lens finding methods. In this paper  we have
shown how \SW provides a lens candidate detection service, crowd-sourcing the
time-consuming work of visually inspecting astronomical images for
gravitationally-lensed features. We invite survey teams searching for lenses in
their wide-field imaging data to contact us if they would like our help.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  ACKNOWLEDGMENTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Acknowledgements}

\input{acknowledgments.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  APPENDICES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Probabilistic Classification Analysis}
\label{appendix:swap}

Our aim is to enable the construction of a sample of good lens candidates.
Since we aspire to making logical  decisions, we define a  ``good candidate''
as one which has a high posterior probability of being a lens, given the data:
$\pr(\LENS|\data)$. Our problem is to approximate this probability. The data~$\data$
in our case are the pixel values of a colour image. However, we can greatly
compress these complex, noisy sets of data by asking each volunteer what they
think about them. A complete  classification in \SW consists of a set of
Marker positions, or none at all. The null set encodes the statement from
the volunteer that the image in question is $\saidNOT$ a lens, while the
placement of any  Markers indicates that the volunteer considers this image to
contain a $\saidLENS$.  We simplify the problem by only using the Marker
positions to assess whether the volunteer  correctly assigned the
classification $\saidLENS$ or $\saidNOT$ after viewing (blindly) a member of
the training set of subjects.

How should we model these compressed data? The circumstances of each
classification are quite complex, as are the human classifiers in general: the
volunteers learn more about the problem as they go, but also inevitably make
occasional mistakes (especially when classifying at high speed).
To cope with this uncertainty, we assign a
simple software {\it agent} to partner each volunteer. The agent's task is to
interpret their volunteer's classification data as best it can, using a model
that makes a number of necessary approximations. These interpretations will
then include uncertainty arising as a result of the volunteer's efforts and
also the agent's approximations, but they will have two important redeeming
features. First, the interpretations will be quantitative (where before they
were qualititative),  and thus will be useful in decision-making. Second, the
agent will be able to predict, using its model, the probability of a test
subject being a $\LENS$, given both its volunteer's classification, and its
volunteer's experience. In this appendix we describe how these agents work,
and other aspects of the \SW analysis pipeline (SWAP).

\subsection{Agents and their Confusion Matrices}
\label{appendix:swap:probabilities}

Each agent assumes that the probability of a volunteer recognising any given
simulated lens as a lens is some number, $\pr(\saidLENS|\LENS,\training)$, that
depends only on what the volunteer is currently looking at, and all the
previous training subjects they have seen (and not on what type of lens it is,
how faint it is, what time it is, \etc). Likewise, it also assumes that the
probability of a volunteer recognising any given dud image as a dud is some
other number, $\pr(\saidNOT|\NOT,\training)$, that also depends only on what the volunteer is currently looking at, and all the
previous training subjects they have seen. These two probabilities define a
2 by 2 ``confusion matrix,'' which the agent updates, every time a
volunteer classifies a training subject, using the following
very simple estimate:
\be
  \pr(``X"|X,\training) \approx \frac{N_{``X"}}{N_X}.
  \label{eq:app:fraction}
\ee
Here, $X$ stands for the true classification of the subject, \ie either
$\LENS$ or $\NOT$, while $``X''$ is the corresponding classification
made by the volunteer on viewing the subject. $N_X$ is the number of
lenses the volunteer has been shown, while $N_{``X"}$ is the number of
times the volunteer got their classifications of this type of training subject
right. $\training$ stands for all
$N_{\LENS} + N_{\NOT}$ training data that the agent has heard about to
date.

The full confusion matrix of the $k^{\rm th}$ volunteer's agent is therefore:
\begin{align}
  \CM^k &=
  \begin{bmatrix}
    \pr(\saidLENS|\NOT,\trainingk) & \pr(\saidLENS|\LENS,\trainingk) \\
    \pr(\saidNOT |\NOT,\trainingk) & \pr(\saidNOT |\LENS,\trainingk)
  \end{bmatrix}, \notag \\
        &=
  \begin{bmatrix}
    \CM_{LN} & \CM_{LL} \\
    \CM_{NN} & \CM_{NL}
  \end{bmatrix}^k.
  \label{eq:confmat}
\end{align}
Note that these probabilities are normalized, such that
$\pr(\saidNOT |\NOT) = 1 - \pr(\saidLENS|\NOT)$.

Now, when this volunteer views a test subject,
it is this confusion matrix that will allow their agent to update the
probability of that test subject being a $\LENS$. Let us suppose that
this subject has never been seen before: the agent assigns a
prior probability that it is (or contains) a lens is
\be
  \pr(\LENS) = p_0
\ee
where we have to assign a value for $p_0$. In the \cfhtls, we might expect
something like 100 lenses in 430,000 images, so $p_0 = 2\times10^{-4}$
is a reasonable estimate. The volunteer then makes a classification $C_k$
($= \saidLENS$ or $\saidNOT$).
We can apply Bayes' Theorem to derive how the agent should
update this prior probability into a posterior one using this new information:
\begin{align}
  \label{eq:app:first}
  & \pr(\LENS|C_k,\trainingk) = \\
  & \frac{\pr(C_k|\LENS,\trainingk)\cdot\pr(\LENS)}
{\left[ \pr(C_k|\LENS,\trainingk)\cdot\pr(\LENS) + \pr(C_k|\NOT,\trainingk)\cdot\pr(\NOT) \right]},
  \notag
\end{align}
which can be evaluated numerically using the elements of the confusion
matrix.

\subsection{Examples}
\label{appendix:swap:examples}

Suppose we have a volunteer who is always right about the true
nature of a training subject.
Their agent's confusion matrix would be
\be
  \CM^{\rm perfect} =
  \begin{bmatrix}
    0.0 & 1.0 \\
    1.0 & 0.0
  \end{bmatrix}.
\ee
On being given a fresh subject that actually is a $\LENS$, this hypothetical
volunteer would submit $C = \saidLENS$.  Their agent would then calculate the
posterior probability for the subject being a $LENS$ to be
\begin{align}
  \pr(\LENS|\saidLENS,\trainingk) &= \frac{1.0 \cdot p_0}
           {\left[ 1.0\cdot p_0 + 0.0\cdot(1 - p_0) \right]}
   &= 1.0,
\end{align}
as we might expect for such a {\it perfect} classifier.  Meanwhile, a
hypothetical volunteer who (for some reason) wilfully always submits the wrong
classification would have an agent with the column-swapped confusion matrix
\be
  \CM^{\rm obtuse} =
  \begin{bmatrix}
    1.0 & 0.0 \\
    0.0 & 1.0
  \end{bmatrix},
\ee
and would submit $C = \saidNOT$ for this subject. However, such a volunteer
would nevertheless be submitting useful information, since given the above
confusion matrix, their agent would calculate
\begin{align}
  \pr(\LENS|\saidNOT,T_k) &= \frac{1.0 \cdot p_0}
           {\left[ 1.0\cdot p_0 + 0.0\cdot(1 - p_0) \right]}
   &= 1.0.
\end{align}
{\it Obtuse} classifiers turn out to be as helpful as {\it perfect} ones, because
the agents know to trust the opposite of their classifications.

\subsection{Online SWAP: Updating the Subject Probabilities}
\label{appendix:swap:examples}

Suppose the $k+1^{\rm th}$ volunteer now submits a classification, on the same
subject just classified by the $k^{\rm th}$ volunteer. We can generalise
\Eref{eq:app:first} by replacing the prior probability with the current
posterior probability:
\begin{align}
  \label{eq:app:update}
  \pr(\LENS & |C_{k+1},\training_{k+1},\data) = \\
  & \frac{1}{Z} \pr(C_{k+1}|\LENS,\training_{k+1}) \cdot \pr(\LENS|\data) \\ \notag
{\rm where}\;\; Z = & \pr(C_{k+1}|\LENS,\training_{k+1})\cdot\pr(\LENS|\data) \\ \notag
      & + \pr(C_{k+1}|\NOT,\training_{k+1})\cdot\pr(\NOT|\data), \notag
\end{align}
and $\data = \{C_k,\trainingk\}$ is the set of all previous
classifications, and the set of training subjects seen by each of those
volunteers.
$\pr(\LENS|\data)$ is the fundamental property of each test subject that
we are trying to infer. We track $\pr(\LENS|\data)$ as a function of time,
and by comparing it to a lower or upper thresholds, make decisions about
whether to retire the subject from the classification interface or
promote it in \Talk, respectively.

\subsection{Information Gain per Classification, Agent ``Skill'' and ``Contribution''}
\label{appendix:swap:examples}

With an agent's confusion matrix in hand we can compute the
\emph{information} generated in any given classification. This will
depend on the confusion matrix elements (\Eref{eq:confmat}) but also on
the probability of the subject being classified containing a lens. The
quantity of interest is the relative entropy, or Kullback-Leiber
divergence, between the prior and
posterior probabilities for the possible truths $T$
given the submitted classification $C$:
\begin{align}
\information =& \sum_T \pr(T|C) \log_2 \frac{\pr(T|C)}{\pr(T)}     \notag \\
             =&    \pr(\LENS|C) \log_2 \frac{\pr(C|\LENS)}{\pr(C)} \notag \\
             +&    \pr(\NOT|C)  \log_2 \frac{\pr(C|\NOT)}{\pr(C)},
\end{align}
where, as above, $C$ can take the values $\saidLENS$ or $\saidNOT$.
Substituting for the posterior probabilities using \Eref{eq:app:first} we get
an expression that just depends on the elements of the
confusion matrix $\CM$ and the pre-classification subject lens
probability $\pr(\LENS) = p$:
\begin{align}
\information =    &     p \frac{\CM_{CL}}{p_c} \log_2 \frac{\CM_{CL}}{p_c}  \notag \\
                  & +(1-p)\frac{\CM_{CN}}{p_c} \log_2 \frac{\CM_{CN}}{p_c},
  \label{eq:app:infogain}
\end{align}
where the common denominator $p_c = p\CM_{CL} + (1-p)\CM_{CN}$. This
expression has many interesting features.  If $p$ is either zero or one,
$\information(C) = 0$  regardless of the value of $C$ or the values of
the confusion matrix elements: if we know the subject's status with
certainty, additional classifications supply no new information. If we
set $p$ to be the prior probability, \Eref{eq:app:infogain} tells us how
much information is generated by classifying it all the way to $p = 1$
(which a perfect classifier, with $\CM_{LL} = \CM_{NN} = 1$, can do in a
single classification). For a prior probability of $2\times 10^{-4}$,
12.3 bits are generated in such a ``detection.''  Conversely, only
0.0003 bits are generated during the rejection of a subject with the
same prior: we are already fairly sure that each subject does not
contain a lens! Imperfect classifiers (with $\CM_{LL}$ and $\CM_{NN}$
both less than 1)  generate less than these maximum amounts of
information each classification; the only classifiers that generate zero
information are those that have $\CM_{LL} = 1 - \CM_{NN}$ (or
equivalently, $\CM_{CL} = \CM_{CN}$ for all values of $C$). We might
label such classifiers as ``random'', since they are as likely to
classify a subject as a $\saidLENS$ no matter the true content of that
subject.

\Eref{eq:app:infogain} suggests a useful information theoretical
definition of the classifier skill  perceived by the agent. At a fixed
value of $p$, we can take the expectation value of the information gain
$\information$ over  the possible classifications that could be made:
\begin{align}
\langle\information\rangle   =& \sum_C \sum_T \pr(T|C) \pr(C) \log_2 \frac{\pr(T|C)}{\pr(T)} \notag \\
         =& - \sum_T \pr(T) \log_2 \pr(T) \notag \\
          & + \sum_C \pr(C) \sum_T \pr(T|C) \log_2 \pr(T|C) \notag \\
         =&         p  \left[ \mathcal{S}(\CM_{LL}) + \mathcal{S}(1-\CM_{LL}) \right] \notag \\
          &     +(1-p) \left[ \mathcal{S}(\CM_{NN}) + \mathcal{S}(1-\CM_{NN}) \right] \notag \\
          & - \mathcal{S}\left[ p    \CM_{LL}       + (1-p)(1-\CM_{NN})     \right] \notag \\
          & - \mathcal{S}\left[ p (1-\CM_{LL})      + (1-p)   \CM_{NN}      \right]
\end{align}
where $\mathcal{S}(x) = x \log_2{x}$. If we choose to
evaluate $\langle\information\rangle$ at $p = 0.5$, the result has some
useful properties. While random classifiers presented with  $p = 0.5$
subjects have $\skill = 0.0$  as expected, perfect classifiers appear to
the agents to have  $\skill = 1.0$. This suggests that  $\skill$, the
amount of information we expect to  gain when a classifier is presented
with a 50-50 subject, is a reasonable quantification of
\emph{normalised skill}. A consequence of this choice is that the
integrated skill (over all agents' histories) should come out to be
approximately
equal to the number of subjects in the survey, when the search is
``complete'' (and all subjects are fully classified). Therefore, a
particular agent's integrated skill is a reasonable
measure of that classifier's
\emph{contribution} to the lens search.

We conservatively initialize both elements of each  agent's confusion
matrix to be $\CM^0_{LL} = \CM^0_{NN} = 0.5$, that of a maximally ambivalent  random classifier, so
that all agents start with zero skill. While  this makes no allowance
for volunteers that actually do have previous experience of what
gravitational lenses look like, we might expect it to help mitigate
against false positives. Anyone who classifies more than one image (by
progressing beyond the tutorial) makes a non-zero information
contribution to the project.

The total information generated during the \cfhtls project is shown in
\Tref{tab:crowd:contributions}. Interpreting these numbers is not easy, but we
might do the following. Dividing this by the amount of information it takes to
classify a \SW subject all the way to the detection threshold (lens
probability 0.95), and then multiplying by the survey inefficiency gives us a
very rough estimate for the effective number of detections corresponding to
the crowd's contribution: these are 2830 and 25 bits for \StageOne and \StageTwo
respectively.  These figures are close to the numbers of detections given in
column 7 of the table.
% The uncertainty in the interpretation of the total
% information generated provides some justification for our focus on the
% expected information gain as a measure of volunteer contribution.

\subsection{Uncertainty in the Agent Confusion Matrices}
\label{appendix:swap:uncertainty}

Finally, the confusion matrix obtained from the application of
\Eref{eq:app:fraction} has some inherent noise which reduces as the
number of training subjects classified by the agent's volunteer
increases. For simplicity, the discussion has thus far assumed the case
when the confusion matrix is known perfectly; in practice, we allow for
uncertainty in the agent confusion matrices by averaging over a small
number of samples drawn from Binomial distributions characterised by the
matrix elements $\pr(C_k|\LENS,\trainingk)$ and  $\pr(C_k|\NOT,\trainingk)$. The
associated standard deviation in the estimated subject probability
provides an error bar for this quantity.

% For ease of notation, we will denote $\pr(C_k|\LENS,T_k)\equiv p_L$ and
% $\pr(C_k|\NOT,T_k)\equiv p_N$. In reality, there is a probability distribution
% for both $p_L$ and $p_N$. Let $p_0$ be the prior probability of the subject
% being a lens. Then the posterior probability, $p_0'$ of the subject being a
% lens after the classification $C_k$ is
% \be
%   \label{eq:app:sec}
% p_0' = \frac{p_L p_0}{\left[ p_L p_0 + p_N (1- p_0) \right]},
% \ee
% The posterior probability distribution $p_0'$ can be obtained by marginalizing
% over the probability distributions of $p_L$, $p_N$ and the prior probability
% distribution $p_0$ such that,
% \be
% P(p_0') = \int p_0' P(p_L) P(p_N) P(p_0) dp_L dp_N dp_0\,.
% \ee
% This marginalization is not analytically tractable. Therefore, we have
% implemented the following Monte-Carlo solution for this problem.
%

%Finally, we also need to update the confusion matrix of an agent and obtain the
%variance on each element of the matrix, once a training subject has been
%classified. We would like to derive the posterior probability of the
%probability elements $p_1$ and $p_2$ given their prior probabilities. For this
%purpose, we can again make use of Bayes' theorem,
%\begin{equation}
%P(p_x'|N_{"X"},N_X,T) = \frac{P(N_{``X"}|p_x',N_X,T) P(p_x'|N_X,T)}{\sum_{N_{``X"}} P(N_{``X"}|p_x',N_X,T) P(p_x'|N_X,T)}
%\end{equation}
%Here, $P(N_{``X"}|p_x',N_X,T)$ is a binomial distribution, although this is not
%true strictly speaking given that our agents are learning and the values of the
%confusion matrix are moving. Modelling the learning curve of our users is yet
%another complicated extension we could think about.

\subsection{Offline SWAP}
\label{appendix:swap:offline}

The probabilistic model described above does not need to be implemented as an
online inference. Indeed, it might be more appropriate to perform the inference
of all agent confusion matrix elements and Subject probabilities
simultaneously, so that the early classifications are not effectively
downweighted as a result of the agent's ignorance. It might also be that this
ignorance builds in some conservatism to the system, reducing the noise due
the early classifications if they are unreliable. In the joint analysis, the
basic assumption that is built into the agents, that their volunteers
have innate and unchanging talent for lens spotting parameterised by two
constant confusion matrix elements which simply need to be inferred given the
data, is implemented in full. The effect is that of applying the time-averaged
confusion matrices, rather than one that evolves as the agents (and in the
real world, the volunteers) learn.

The mathematics of the offline inference are presented elsewhere (in
preparation). Here we briefly note that we maximize the joint posterior
probability distribution for all the model parameters (some 66,000 confusion
matrix elements and 430,000 subject probabilities) with a simple
expectation-maximisation algorithm. This procedure takes approximately the
same CPU time as the \StageTwo online analysis, because no matrix inversions are
required in the algorithm. The algorithm scales well and is actually faster
than the online analysis with the larger \StageOne dataset. The
expectation-maximisation algorithm is robust to initial starting parameters
in, e.g., initial agent confusion matrix elements and Subject probabilities.
The plots in \Sref{sec:results:sample} show the difference in performance
between the online and offline analyses.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  REFERENCES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% MNRAS does not use bibtex, input .bbl file instead.
% Generate this in the makefile using bubble script in scriptutils:

% bubble -f paper-lcr.tex references.bib
% \input{paper-lcr.bbl}

\bibliographystyle{apj}
\bibliography{references}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\label{lastpage}
\bsp

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%