### 2015-04-20 doc/addresses.tex
Correcting Edwin's affiliation

\def\oxford{Dept.\ of Physics, University of Oxford, Keble Road, Oxford, OX1 3RH, UK}
\def\oxfordeng{Dept.\ of Engineering Science, University of Oxford, Parks Road, Oxford, OX1 3PJ, UK}
\def\kipac{Kavli Institute for Particle Astrophysics and Cosmology, Stanford University, 452 Lomita Mall, Stanford, CA 94035, USA}
\def\ipmu{Kavli IPMU (WPI), University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa 277-8583, Japan}
\def\zooniverse{Zooniverse, c/o Astrophysics Department, University of Oxford, Oxford OX1 3RH, UK}
\def\adler{Adler Planetarium, Chicago, IL, USA}
\def\lausanne{EPFL, Lausanne, Switzerland}
\def\zurich{Department of Physics, University of Zurich, Switzerland}
\def\paris{Institut d’Astrophysique de Paris, UMR7095 CNRS – Universit\'e Pierre et Marie Curie, 98bis bd Arago, 75014 Paris, France}
\def\icg{Institute of Cosmology and Gravitation, University of Portsmouth, Dennis Sciama Building, Portsmouth P01 3FX, UK}

\def\pjmemail{\tt pjm@slac.stanford.edu}
\def\amemail{\tt anupreeta.more@ipmu.jp}