### 2018-05-29 doc/sw-guidelines-policy-HSC.tex
HSC Guidelines added

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Space Warps Guidelines & Policy
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[a4paper,twocolumn]{article}

% Packages:
\usepackage{xspace}
\usepackage{verbatim}
\usepackage{hyperref}
\usepackage{enumitem}

% Macros:
\input{macros.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{Guidelines and Policy for the HSC Strong Lensing Science Team using \SW}
\author{Aprajita Verma, Phil Marshall \& Anupreeta More}
\date{v1: May 29, 2018}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract} 
\noindent \SW was conceived as a strong lens discovery {\it service}
that harnesses the power of citizen scientists. The HSC strong
lensing group (hereafter HSC survey science team or HSC SST) can
make use of \SW within the logistical constraints. In this document,
we discuss how the HSC SST will interact with \SW and the
relationship between the HSC SST and the \SW collaboration.  This
document includes guidelines for proposing projects, data provision,
data policy and for publications of \SW-enabled discoveries.
\end{abstract}

\setcounter{footnote}{0}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction to \sw}

The \SW website is designed to motivate and enable tens of thousands
of people to perform the scientific tasks of strong gravitational lens
identification.  \SW comprises a primary `classification' interface
(\texttt{http://spacewarps.org}) that allows citizen scientists to
inspect images and mark those that contain potential gravitational
lenses. A mandatory part of this inspection is the classification of
simulated lenses (appropriate for the survey being conducted). The
classification of simulated lenses (and pre-inspecetd blank or false positives) acts as an interactive training
activity that also allows us to establish the citizens' likelihood of
identifying potential lenses and blank fields. This is fed into the
\SW Analysis Pipeline (SWAP) that analyses the classifications made
per image (or subject in Zooniverse terms) giving a probability of
whether an image contains a lens candidate or not.

A key part of the \SW project is the ability for all members of the
\SW collaboration (see Section \ref{sec:members}) to communicate with
each other and discuss interesting candidates through
\Talk\footnote{\Talk (e.g. \texttt{https://www.zooniverse.org/projects/aprajita/space-warps-hsc/talk}) is
  a forum for \SW collaborators to discuss individual sources and the
  \SW project in general. Bugs can also be reported. This is the
  direct portal for communication with the \SW community.}.

In order to use the \SW interface, the HSC Survey Science Team
\textbf{must define and propose a project to the \SW PIs} who will
discuss whether the project is suitable with the HSC SST. The \SW PIs will
then liaise further with the SST to implement the project in the \SW
interface. Details are given in Section \ref{sec:project}.

\vspace{4mm} The PIs  and co-founders of the \SW project are the authors of this
document, P.\,Marshall, A.\,More and A.\,Verma, please do not hesitate
to contact us if you require any further information or have any
questions. The PIs may be collectively contacted via the email address
\href{mailto:spacewarpspi@googlegroups.com}{spacewarpspi@googlegroups.com}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Definition of collaboration member types}
\label{sec:members}

In this document, we will refer to the entirety of the HSC strong
lensing science group as the HSC Survey Science Team (HSC SST) to
encompass all of the HSC strong lensing community.

We now discuss the relationship between the \SW collaboration and the
HSC SST, who wish to use the \SW interface to investigate their
data. The HSC SSTs will liaise with the \SW science and development
teams to ingest their survey data into the \SW interface. There may be
some overlap between current \SW members and SSTs.

When we refer to the \SW collaboration, it comprises the \SW science
team, \SW development team, and all the volunteers who have logged
in\footnote{with a Zooniverse ID. It is possible to classify
  anonymously, and we count those classification but as we don't know
  who they are, we are unable to collaborate with them directly.} and
contributed classifications to the project.

If required, an expanded definition of the \SW membership is given in
the document ``\SW Publication Policy: CFHT-LS''\footnote{available
  from
  \texttt{\url{https://github.com/drphilmarshall/SpaceWarps/blob/master/doc/sw-publication-policy-cfhtls.pdf}}},
however this will not be relevant for \SW-HSC publications, see
Section \ref{sec:papers} for details.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Defining a suitable \sw science project}
\label{sec:project}

The HSC SST is expected to propose a project to the \SW collaboration
that can be best or only be done using the power of citizen
scientists. Typically, this involves visual inspection tasks that are
too time consuming for the SST themselves to execute. As such, any
science programme conducted through \SW should be complementary, not
in competition, to the lens finding activities being carried out by
the SST.

For example, in the first \SW project, the collaboration conducted a
lens finding search with \SW over the entire CFHT-LS survey area. The
goal was to increase the completeness of strong lenses in the survey
areas. Both RingFinder and ArcFinder algorithms had been used to
discover lenses of different types. \SW complemented these searches
with the potential to find strong lenses missed by these
algorithms. The method, proposed by Anupreeta More (who served as the
SST for this project), was to inspect the entire survey area by
getting citizens to view $\sim$430,000 tiles of the survey area. This
was a \textit{blind} search. Such a search could not be conducted by a
normal sized SST.

\textit{Targeted} searches are also possible, as we performed in the
second \SW project on the VICS82 survey area. Here we inspected
$\sim$40,000 images from the VICS82 survey centred on likely strong
lens targets, e.g. LRGs, groups, clusters, quasars.  Again, this was
more images than the VICS82 SST could manage to inspect alone.

Another approach is to use the \SW service to inspect the normally
large candidate lists generated by automated algorithms or searches,
which are typically dominated by false positives.

\subsection{HSC specific thoughts}

\begin{enumerate}
\item The \SW facility should be used for science projects that cannot be completed by the Strong Lens Group itself. 
\item Any \SW search should be complementary, not in competition, to the lens finding being conducted by the HSC SST.
\item While \SW uses citizen science, its primary goal is \textbf{not outreach}, it is harnessing citizen power for visual classifications (and eventually modelling) to perform science.
\item A duplicate project is not encouraged unless there is sufficient scientific motivation and justification to do so.
\item \SW requires volunteers/citizens to give up their personal time to conduct the experiment, this should not be taken lightly, the crowd are motivated by contributing to real science not just doing things for the sake of them. This is the reason the suitability of the project for citizen scientists should be carefully considered. The \SW PIs will be able to discuss any planned \SW-HSC projects with the HSC SST. The crowd needs to be engaged and feel like they are contributing while the entire \SW-HSC project runs.
\item The success of \SW-HSC will rely on input from the HSC SST working with the \SW collaboration, dedicated effort is required for this and the HSC SST should identify team members who can commit time to the \SW-HSC tasks for this before embarking on a project. The \SW PIs will help to establish the main efforts required, these will depend on the actual project put forward by the HSC SST (e.g.\ image preparation, simulations, interacting with volunteers on \Talk).
\end{enumerate}

\textbf{As such, \SW should be considered to be an integral/complementary part of the HSC SST lens finding strategy, rather than an alternative method.}

The scientific project, process and outcome is the domain of the
SST. The \SW collaboration enables this work by making it a live
project on the website. We do not, per se, expect the \SW
collaboration lead the resultant papers, we just request some of us
are included on discovery papers in return for our help, please see
Section \ref{sec:papers} for more details on the publication policy.

The goal of the \SW PIs and the \SW collaboration is to enable the
HSC SST's science project in \sw and make it work. Using the
automatic analysis pipeline (SWAP), the \SW PIs we will send the HSC
SST lists of the candidates selected by the citizens (at some regular
intervals or at the end of a project), we will then discuss strategies
for further analysis based on our experience of previous
projects. Some members of the \SW PIs and team may become more
involved in analysis steps such as grading of the \sw candidates or
assist with any scientific investigations and follow-up of the
candidates, at the HSC SST's request.

\subsection{Possible \SW-HSC survey strategies}

In our experience, a citizen science project like SW cannot function
successfully without significant support and co-ordination of the
SST. Thus, we recommend that all HSC members with an interest in
strong lensing are told about \SW and how it can complement their lens
finding effort. Our goal is to enable your science.

We also encourage the HSC SST to consider the timing for when they
wish to ingest the images into \SW so that it is in line with the
interests of the HSC SST and those of the HSC consortium.

Possible projects for the HSC SST could consider (just to get a flavour of what can be done)
\begin{itemize}
\item \textbf{New Data}: Inspection of areas (targeted or blind) that the SST has not had time to inspect, i.e. use the crowd to increase the speed at which potential lens candidates are discovered
\item \textbf{Comparison of Expert vs Citizen Classification}: HSC SST and citizen's view images in parallel (targeted or blind)
\item \textbf{Completeness of Expert Classification}: Citizen's classify images already inspected by the HSC SST (targeted or blind)
\item \textbf{Helping the robots}: Use citizens to weed out false positives from candidate samples generated by algorithm-based or machine learning searches
\item \textbf{Training sets}: Blind inspection to generate a large training set for machine learning algorithms.
\end{itemize}

Note that the projects that are proposed to \SW need not be one of the
above. Multiple projects could also be  run in a mixture or phased set
of experiments. Please do consider the HSC timeline with reference to
data release and the searches the HSC SST want to make themselves.

If the HSC SST wants to inspect targets within the consortium and is
concerned about being scooped by the crowd, please be assured that
there is little scope or motivation to do this. If the project is well
defined, the goal of \SW-HSC will be a complementary part of the HSC
SST lens finding strategy. By definition this means that \SW is
working with the HSC SST as it is designed to do. In addition, the
images could be stripped of WCS info (described in Section
\ref{sec:data}), meaning that no-one other than the SST can follow-up
or catalogue the targets\footnote{We cannot exclude the possibility
  that individual sources might be serendipitously identified e.g. by
  recognisable morphologies.}.  The HSC SST could also perform their
inspection before the images are passed to \SW. Any lens candidates
discovered by the SST would then be labelled as `known lens
candidates' in \sw\,\Talk. This applies to candidates discovered
through visual inspection or algorithm based searches.

\subsection{What to do next: Project Proposals}
Once the HSC SST have narrowed down what search to be conducted \SW,
please approach the \SW PIs (Phil Marshall, Anupreeta More, Aprajita
Verma, collectively contacted via
\href{mailto:spacewarpspi@googlegroups.com}{spacewarpspi@googlegroups.com})
with a short description of the project that includes the following
information.

\begin{itemize}
\item Survey area
\item Bands that could be included
\item Total number of images to be viewed (approx.)
\item Targeted survey or blind field search\footnote{Targeted: inspection of a list of candidates; Blind: inspection of tiles of a survey area}
\item Brief outline of the overall goal(s) of that inspection, highlighting how this is complementary (not competing) with the goals of the HSC lens finding group
\item Proprietary data issues 
\item Identify an initial core team who will lead and work closely with \SW
\end{itemize}

The \SW PIs will then arrange a telecon with the HSC SST to discuss
the project further. This includes discussing the initial effort
required and roles for the the SST team for expected tasks.

The original \SW papers, by definition, provide a complete description of the system, and its results from the CFHT-LS survey. The SST will be able
to read these papers before designing their own \SW project, and cite them as justification for some of their experimental decisions. 
See \href{https://github.com/drphilmarshall/SpaceWarps/raw/master/doc/sw-system-published.pdf}{Paper I} \& \href{https://github.com/drphilmarshall/SpaceWarps/raw/master/doc/sw-cfhtls-published.pdf}{Paper II}.

\subsection{What to do next: Example Timeline}
The SST is expected to interact with \SW in a timely manner with the following rough guidelines in collaboration with \SW team
\begin{itemize}
\item at least 6 months before launch: determine project goals
\item 2-4 months before launch: select training images and write appropriate feedback messages
\item 2 months before launch: deliver and ingest subject images 
\item 1-2 month before launch: update website content with survey specific text and images
\item 1 month before launch: check inspection data batches and determine retirement strategy
\item 1 month before launch: revisit survey strategy plans, implementation with \SW team
\item 2 weeks before launch: run a beta test (interface, data sets, training frequency, SWAP), all hands on deck for interaction on talk
\item Launch and subsequently while project is running: all hands on deck for interaction on talk, blog posts, interaction with \SW PIs and zooniverse
\end{itemize}
see also Section \ref{sec:data}.

\hspace{6mm}

In the remaining sections we look at discovery publications that might
arise from future \SW projects, and suggest reasonable guidelines for
deciding on their authorship. First, however, we remind ourselves about data
access via \SW.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Survey Data Policy}
\label{sec:data}

\SW is a public website. All images displayed there are by definition in the
public domain, and so must be expected to be downloaded, copied and
redistributed by any \SW user. This is a \textbf{good thing}: images of the sky taken
with publicly-funded telescopes belong to everyone.

However, some surveys, including HSC, come with their own proprietary
access policy. It is the responsibility of the HSC SST to provide
images to \SW in a way that is consistent with their own data access
rules. There are two things that SSTs can do to in order to respect
any proprietary period that they have imposed on themselves:
\begin{enumerate}

\item Add a ``LICENSE'' keyword to the FITS and PNG image headers, explaining
what the rules for redistribution of these images are. This will almost
certainly be ignored, but it would be a nice reminder that nothing comes free
of either cost or responsibility. Other keywords could also be included as
well: data provenance is important, and links to useful survey webpages would
be most welcome!

\item Remove all WCS information from all images provided. The \SW interface ``dashboards'' do allow FITS and JPG files to be downloaded. If images from a proprietary dataset being made available is a concern for an SST, they should remove the WCS. This also has the consequence that it minimises the potential ``scooping'' of  lens candidates by anyone other than the SST. 
In practice, the objects contained in the \SW images will be too faint
for anyone without access to a large telescope to observe, and may also be
absent from any public catalogs, so the opportunities for follow-up will be
quite limited. Images
with field of view less than about 3~arcminutes in diameter are not solvable
by \texttt{astrometry.net}. \textbf{Please note}, however, that we cannot exclude the possibility that object co-ordinates are posted in \Talk. For example, co-ordinates of previously published lens candidates or recognisable fields from CFHT-LS have been posted in \Talk by a very small number of citizens.

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Data Preparation \& the \SW interface}
\label{sec:data}

\subsection{Collaborating with \SW Science Team}

Implementing a new project based on different survey data will require support from the \SW consortium, in four respects:

\begin{enumerate}

\item In {\bf preparing data} for the site, and uploading it onto the
Zooniverse servers. There is a reasonable amount of work in preparing data and images by the HSC SST for delivery to \SW. The \SW team will provide assistance to the HSC SST in data preparation. The benefits of using the \SW service will far outweigh the time needed to prepare data.  The \SW Science Team can advise on suitable Training
Subjects (images of simulations and false positives) and image display settings, and also on formatting the data
ready for ingestion and display. A mandatory element of the training subjects is the production of large samples of simulated lenses. \SW will provide resources to assist generating these simulations.

\item In {\bf reconfiguring the site itself} ready for the new data. The HSC
SST will need to include new Spotter's Guide images and text,
modified tutorial content, and additional survey-specific site content;
the \SW Science Team can help with all of this.

\item In {\bf maintaining the site} so that it continues to function and deliver
correctly. This will involve technical support from the Zooniverse and \SW science teams,
but will also require the continued engagement of the wider
collaboration by the SST in \Talk. Since this social filter is
a key part of the sample refinement, the latter may involve a
significant time commitment.

\item In {\bf running the \SW analysis pipeline} (SWAP) which will be the \sw PIs responsibility. This will provide citizen selected candidates to be sent to the HSC SST.

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Resulting publications for \SW-HSC }
\label{sec:series}

\subsection{Journal Papers}
\label{sec:papers}
In return for the work outlined in Section \ref{sec:data} in setting up the new surveys, the \SW PIs will nominate a core \SW collaboration members as authors on \textbf{papers that
present discovery of a lens or sample of lenses enabled by the \SW system}. A ``discovery'' is defined by the announcement of the source(s) to the astronomical community in a journal paper. We would expect the \SW author list to be a small number (approx. ten), primarily the Principal Investigators, a few others working on the ingestion of the new data or moderating \Talk. All post-discovery follow-up papers would of course be exempt from this (unless individuals from \sw team had contributed sufficiently such that the lead author wished to include them).

We request that authors of \SW-enabled discovery papers should circulate drafts to the list of \SW authors and, if possible, allow a minimum of two weeks to give time for comments. We also request that any papers based on \SW enabled discoveries should cite the \SW system papers. This policy document will be updated with the correct citations to the \SW system papers when available.

\SW runs an analysis pipeline (SWAP) that will be tailored to the science goals of each new survey. Any large changes to the analysis code will be documented  by the \SW Science Team. The \SW Science Team reserve the right to update the \SW methodology papers.

\subsection{Zooniverse Communications}
\label{sec:comm}

Any \SW collaboration member, simply by virtue of their Zooniverse
registration, may write a Zooniverse
blog post or \Talk reports describing their
investigation of any lens candidate they find in \SW. This, along with
posts in  \Talk, is the primary means by which we expect \SW-HSC collaboration
members (citizens participating in the project) will communicate their findings to the rest of the
astronomical community. The investigation of any \SW images that are
provided without world coordinate system (WCS) information will be
necessarily limited but the \SW Zooniverse \Talk and blog posts will contribute to the collective knowledge and understanding of the presented system(s).

As \SW classifications are a community wide activity, it is impossible to attribute the discovery of a candidate to a single community member, or group  of members. Therefore, in recognition of their contribution to the \SW project, \SW community members will be listed (on their approval) on the \SW members web page. This web page will be linked in every ensuing \SW publication. If a community member makes a significant contribution e.g.\ in the further investigation of a lens candidate with modelling tools, they are strongly encouraged to write Zooniverse posts. This may lead to them being invited to join journal publications by the lead author of a ``discovery'' publication.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Summary of Publication Guidelines}
\label{sec:publ}
Any \SW series and \SW enabled discovery papers should include
\begin{itemize}
\item citations to the \SW system papers, \SW I \& II
\item a small number of \SW authors proposed by the \SW PIs (see \ref{sec:papers}).
\item the collaboration should be acknowledged in the acknowledgement section with a link to the collaboration membership page (e.g. \href{https://spacewarps.org/#/projects/CFHTLS/contributors}{List of Contributors to CFHT-LS}).
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%